---
title: "01_json_transform"
format: html
editor: visual
---

## json transformation

Make json files to tables for easier work with R graphics etc.

```{r}
library(jsonlite)
# library(tidyjson)
library(tidyverse)
```

### formulas

```{r}
x <- jsonlite::fromJSON("../data/formulas.json", simplifyVector = F)

length(x) # number of languages

# str(x[1])
str(x[[2]][1])
```

```{r, eval = FALSE}
# test run with one language
df <- tibble(
  doc_key = names(x$de),
  triplets = x$de
) %>% 
  unnest_longer(
    triplets,
    indices_include = TRUE,
    names_repair = "minimal"
  ) %>% 
  rename(triplet_id = triplets_id) %>% 
  mutate(
    source_id = map_chr(triplets, 1),
    target_id = map_chr(triplets, 2),
    text = map_chr(triplets, 3),
    author_name = map_chr(triplets, 4),
    year_birth = map_dbl(triplets, 5, .default = NA), # need to deal with NULL
    year_death = map_dbl(triplets, 6, .default = NA)
  ) %>% 
  select(doc_key, triplet_id,
         source_id, 
         target_id, 
         text, 
         author_name, 
         year_birth, year_death)

df
```

```{r}
df_binded <- tibble()
temp <- NULL # keep extracted language data
l <- NULL # write language tag

for (i in 1:length(x)) {
  # select language corpus object
  temp <- x[i]
  # store language label
  l <- names(temp[1])
  
  df <- tibble(
    # extract document title from list name
    doc_key = names(temp[[1]]),
    triplets = temp[[1]] # extract the list with 3 el data
    ) %>% 
    # unnest list to each element = 1 row
    unnest_longer(
      triplets,
      indices_include = TRUE,
      names_repair = "minimal"
    ) %>% # now we have the main array expanded, go inside
    rename(triplet_id = triplets_id) %>% 
    mutate(
      # paste each inner list element to the respective column
      source_id = map_chr(triplets, 1),
      target_id = map_chr(triplets, 2),
      text = map_chr(triplets, 3),
      author_name = map_chr(triplets, 4),
      year_birth = map_dbl(triplets, 5, .default = NA),
      year_death = map_dbl(triplets, 6, .default = NA)
    ) %>% 
    mutate(lang = l) %>% # add language column & clean
    select(lang, doc_key, triplet_id,
           source_id, 
           target_id, 
           text, 
           author_name,
           year_birth, year_death) 

  # attach to the main table
  df_binded <- rbind(df_binded, df)
  
}

df_binded %>% 
  sample_n(10)
```

```{r}
df_binded %>% 
  filter(lang == "ru") %>% 
  head(20)
```

```{r}
df_binded %>% 
  count(lang)
```

### wikidata

```{r}
w <- fromJSON("../data/wikidata.json")

# head(w)
length(w)

w[[1]][1] # label - england
w[[1]][2] # type -- country
# w[[1]][3] # labels -- long list of spellings
w[[1]][4] # lat -- latitude
w[[1]][5] # lon longitude
w[[1]][6] # instances 

w[[13]][1]
```

```{r}
wiki_df <- tibble(
  entity_id = names(w), # keep id as chr vec, "Q21"
  json_list = w # add json list with data
) %>% 
  mutate(
    label = map_chr(json_list, 1, .default = NA),
    type = map_chr(json_list, 2, .default = NA),
    latitude = map_dbl(json_list, 4, .default = NA),
    longitude = map_dbl(json_list, 5, .default = NA),
    instances = map(json_list, 6, .default = NA) # this is a json array, so turn it to a named list and flatten later
    ) %>% 
  select(-json_list) %>% 
  mutate(instances = map_chr(instances, ~paste(.x, collapse = ", "))) # paste as chr
```

```{r}
glimpse(wiki_df)

unique(wiki_df$type) # default -- city ??

wiki_df %>% 
  filter(type == "default") %>% 
  sample_n(10)
```

### merge

```{r}
glimpse(df_binded) # formulas main table
glimpse(wiki_df) # wikidata simplified
```

```{r}
formulas_df <- df_binded %>% 
  # attach data for FROM
  left_join(wiki_df %>% 
              rename(source_id = entity_id) %>% 
              select(-instances), 
            by = "source_id") %>% 
  rename(from_id = source_id,
         from_placename = label,
         from_type = type,
         from_latitude = latitude,
         from_longitude = longitude) %>% 
  # attach data for TO
  left_join(wiki_df %>% 
              rename(target_id = entity_id) %>% 
              select(-instances),
            by = "target_id") %>% 
  rename(to_id = target_id,
         to_placename = label,
         to_type = type,
         to_latitude = latitude,
         to_longitude = longitude) %>% 
  select(-triplet_id)

glimpse(formulas_df)
```

write

```{r}
write.csv(formulas_df, "../data/formulas_table.csv", row.names = F)
```

```{r}
# write wiki df separately to use for instances later
write.csv(wiki_df, "../data/wiki_df.csv", row.names = F)
```

## fixes

### work with loc types

```{r, eval = F}
# check

# r <- read.csv("../data/formulas_table.csv")
# 
# glimpse(r)
```

Write a table with unique locations from the formula (to work on types)

```{r, eval = F, include = F, echo = F}
loc_f <- r %>% 
  select(from_id, from_latitude, from_longitude, from_placename, from_type) %>% 
  rename(id = from_id, 
         lat = from_latitude,
         lon = from_longitude,
         placename = from_placename,
         type = from_type)

loc_t <- r %>% 
  select(to_id, to_latitude, to_longitude, to_placename, to_type) %>% 
  rename(id = to_id, 
         lat = to_latitude, 
         lon = to_longitude,
         placename = to_placename,
         type = to_type)

colnames(loc_f) == colnames(loc_t)

locs <- rbind(loc_f, loc_t)

locs <- locs %>% 
  distinct()

# write.csv(locs, "../data/loc_unique.csv")
```

Load a table with tagged location types

```{r}
loc_tag <- read_delim("../data/loc_unique_tagged.csv") %>% 
  select(-`...1`)

glimpse(loc_tag)
```

Clean the types

```{r}
loc_tag <- loc_tag %>% 
  mutate(type_new = ifelse(is.na(type_detail), type, type_detail)) %>% 
  select(-type, -type_detail) %>% 
  rename(type_detail = type_new)

loc_tag %>% 
  count(type_detail, sort = T) %>% head(10)
```

Create larger groups of loc types

```{r, include = F, echo = F}
loc_tag <- loc_tag %>% 
  mutate(type_general = type_detail,
         type_general = str_replace(type_general, "historic region", "region")) %>% 
  mutate(type_general = str_replace(type_general, "^land$", "region")) %>% 
  mutate(type_general = str_replace(type_general, "archipelago", "island")) %>% 
  mutate(type_general = str_replace_all(type_general, 
                                        "castle|fort|^castle ruin$|^historic castle/fort$ ", 
                                        "castle/fort")) %>% 
  mutate(type_general = str_replace_all(type_general, "^castle/fort ruin$", "castle/fort")) %>% 
  mutate(type_general = str_replace_all(type_general, "^historic castle/fort$", "castle/fort")) %>% 
  mutate(type_general = str_replace(type_general, "ciity", "city")) %>% 
  mutate(type_general = str_replace(type_general, "mountain pass", "mountain")) %>% 
  mutate(type_general = ifelse(type_general %in% c("hill", "hills", "hilltop"), 
                               "hill", type_general)) %>% 
  
  mutate(type_general = ifelse(type_general %in% c("point", "prison", "battle",
                                                   "church", "cliff", "college", 
                                                   "pillars (also mythological)", 
                                                   "road", "spring", "rock",
                                                   "tower", "beach"
                                                   ), "default", type_general)) %>% 
  mutate(type_general = str_replace(type_general, "street", "city part"),
         type_general = str_replace(type_general, "sea area", "sea"),
         type_general = str_replace(type_general, "kibbutz", "village"),
         type_general = str_replace(type_general, "state", "region")) 

loc_tag %>% count(type_general, sort = T) %>% head(10)
```

### fix formulas

Read the fixed file

```{r}

fch <- read_delim("../data/working_files/formulas_table_check.csv")
glimpse(fch)
```

Remove rows with errors & leave the clean data

```{r}
unique(fch$error)

formulas <- fch %>% 
  mutate(error = ifelse(is.na(error), "", error)) %>% 
  # leave when not an error
  filter(error %in% c("multiple items order", "")) %>% 
  select(-error)
```

```{r}
glimpse(formulas)
```

Re-attach coordinates and location types

*NB load df_binded (chunk 4) and wiki_df (chunk 7 etc)*

```{r}
# attch author's data back
df_authors <- df_binded %>% 
  select(lang, doc_key, author_name, year_birth, year_death) %>% 
  distinct()


formulas <- formulas %>% 
  left_join(df_authors, by = "doc_key")

# quick check
formulas %>% 
  filter(is.na(author_name))

glimpse(formulas)
```

check Lat-Lon

```{r}
formulas %>% 
  filter(is.na(from_id) | from_id == "")

formulas %>% 
  filter(is.na(to_id) | from_id == "")
```

Load fixed wiki df

```{r}
wiki_df <- read_csv("../data/wiki_df.csv")
glimpse(wiki_df)
```

Attach types

```{r}
# glimpse(loc_tag)

wiki_df <- wiki_df %>% 
  left_join(loc_tag %>% 
              select(id, type_detail, type_general) %>% 
              rename(entity_id = id),
            by = "entity_id") 

glimpse(wiki_df)
```

```{r}
# quick fix: two instances of Q29 in wiki_df
wiki_df %>% 
  count(entity_id, sort = T)

wiki_df %>% distinct() %>% nrow()
nrow(wiki_df)

wiki_df <- wiki_df %>% 
  distinct()
```

```{r}
glimpse(formulas)
```

Attach wiki_df data to the main table

```{r}
formulas_l <- formulas %>% 
  #rename old names
  rename(from_placename_old = from_placename,
         to_placename_old = to_placename) %>% 
  
  # attach data for FROM
  left_join(wiki_df %>% 
              rename(from_id = entity_id) %>% 
              select(-instances), 
            by = "from_id") %>% 
  rename(
         from_placename = label,
         from_type_old = type,
         from_latitude = latitude,
         from_longitude = longitude,
         from_type = type_general,
         from_type_d = type_detail) %>% 
  # attach data for TO
  left_join(wiki_df %>% 
              rename(to_id = entity_id) %>% 
              select(-instances),
            by = "to_id") %>%  
  rename(
         to_placename = label,
         to_type_old = type,
         to_latitude = latitude,
         to_longitude = longitude,
         to_type = type_general,
         to_type_d = type_detail) %>%  
  # fix NA in types
  mutate(from_type = ifelse(is.na(from_type), from_type_old, from_type),
         from_type_d = ifelse(is.na(from_type_d), from_type_old, from_type_d),
         
         to_type = ifelse(is.na(to_type), to_type_old, to_type),
         to_type_d = ifelse(is.na(to_type_d), to_type_old, to_type_d))

glimpse(formulas_l)
```

Check

```{r}
formulas_l %>% 
  filter(is.na(from_latitude) | is.na(from_longitude))

formulas_l %>% 
  filter(is.na(to_latitude) | is.na(to_longitude))

formulas_l %>% 
  filter(from_placename_old != from_placename) %>% 
  select(from_id, to_id, text, from_placename_old, from_placename)
```

Remove (still) missing data & old columns

```{r}
formulas_l <- formulas_l %>% 
  filter(!is.na(from_latitude), !is.na(from_longitude)) %>% 
  filter(!is.na(to_latitude), !is.na(to_longitude)) %>% 
  # remove old columns
  select(-from_placename_old, -to_placename_old,
         -from_type_old, -to_type_old)

# glimpse(formulas_l)
```

final reorder

```{r}
# reoder columns
formulas_fin <- formulas_l %>% 
  select(lang, doc_key, 
                from_id, to_id, text,
                author_name, year_birth, year_death, 
                from_placename, from_latitude, from_longitude, 
                from_type, from_type_d,
                to_placename, to_latitude, to_longitude,
                to_type, to_type_d)

glimpse(formulas_fin)
```

### write

```{r}
write_csv(formulas_fin, "../data/formulas_table.csv")
```
