---
title: "01 geodata exploration"
format: md
editor: visual
---

# Explore directions of from_to formula

```{r, warning=F, include=F}
library(tidyverse)

library(MetBrewer)
library(cowplot)
library(kableExtra)
theme_set(theme_minimal())

library(geosphere) # distHavestine()
```

Read data as a table (flatted & merged json-s)

```{r}
formulas <- read.csv("../data/formulas_table.csv")
glimpse(formulas)
```

Keep only ONE from-to direction per poem

```{r}
# formulas %>% 
#   distinct() %>% nrow()
# 
# nrow(formulas)

formulas <- formulas %>% 
  distinct()

# just removes 2 rows
```

N formulas per language

```{r, include=T, echo=F}
formulas %>% 
  count(lang, sort = T)
```

## Most freq loc vs from_to loc

Look if most most freq locations also appear in from_to formula

```{r, include=T, echo=F}
# load loc freq data
loc_freq <- read.csv("../data/locations.csv")

glimpse(loc_freq)

loc_freq <- loc_freq %>% 
  # add ranks within groups
  group_by(corpus) %>% 
  mutate(rank = row_number()) %>% 
  ungroup() %>% 
  rename(lang = corpus)

f_long_from <- formulas %>% 
  select(lang, from_id) %>% 
  rename(qid = from_id)

f_long <- f_long_from %>% 
  rbind(formulas %>% select(lang, to_id) %>% rename(qid = to_id))

f_long %>% 
  group_by(lang) %>% 
  count(qid, sort = T) %>% 
  mutate(rank_ft = row_number()) %>% 
  ungroup() %>% 
  mutate(id = paste0(lang, "_", qid)) %>% 
  select(-lang, -qid) %>% 
  left_join(loc_freq %>% mutate(id = paste0(lang, "_", qid)), by = "id") %>% 
  filter(rank_ft < 21) %>% 
  arrange(lang) %>% 
  select(-lat, -lon) %>% 
  rename(from_to_count = n, 
         total_count = count, 
         rank_total = rank) %>% 
  select(id, qid, lang, label, rank_ft, rank_total, from_to_count, total_count) %>% 
  filter(!is.na(qid))

rm(f_long, f_long_from, loc_freq)
```

It's a long table, just some observations:

-   Czech:

    -   top in both lists (top-20): Bohemian Forest, Prague, Tatra mountains, Danube, Moravia, Vltava, ROME (!) , Czechia, Elbe, Vyšehrad;

    -   many mountains which is not the case for other corpora (Tatra, Giant Mountains, Ural, Sněžka, Alps)

-   DE:

    -   top-20 loc in both: Vienna, Rhine, Paris, Berlin

    -   not many instances overall (only Vienna, Rhine and Paris apprear \> 2 times in from-to)

-   EN:

    -   England, France, Egypt (!), Nile, River Thames, Rome, Ireland, Spain, India

    -   top-1 is Maine (?)

    -   many frequent distant/non-european locations (china, nile);

-   FR:

    -   Paris, Rome, Kythira (wtf);

    -   top-2 locs are very highly cor within lists, overall cor is non-existent; meaning from-to locs are important locs in general (eg Bordeau, Nice, Louvre Palace); some long-distance things in from-to also – Tiber, Ganges, Memphis, Volga, Cairo

-   IT:

    -   Alps,

    -   too little data (all from-to freq is 2 or 1)

-   RU:

    -   Neva, Moscow, SPb, Volga, Crimea, Msc Kremlin, Caucasus, Kyiv, Rome, Nile, Siberia (21 total rank)

    -   very obvious places mostly within Russia, exceptions: Alps, Caucasus, Kyiv, Rome, Nile, Vistula, Seine;

-   SL:

    -   Soča (from-to count: 3, else is 2 or less)

    -   most locs are some within Sl I guess + mostly single counts in from_to

## Distances

I calculated Haversine (bird/plane flight) distance between the from-\>to coordinates, the distance is in km.

### calculation

```{r, eval=F, include=F}
# test on smaller sample
t <- formulas[1:2,]
t

# geosphare library

lat_london <- 51.5074
lon_london <- -0.1278
lat_paris <- 48.8566
lon_paris <- 2.3522

distHaversine(c(lon_london, lat_london), c(lon_paris, lat_paris)) / 1000 # div1000 gives km

# same for two first row from formulas
distHaversine(c(t$from_longitude[1], t$from_latitude[1]), 
              c(t$to_longitude[1], t$to_latitude[1])) / 1000

# check application to the table:
t %>% 
  rowwise() %>% 
  mutate(dist_haversine = distHaversine(c(from_longitude, from_latitude),
                                        c(to_longitude, to_latitude)) / 1000) %>% 
  ungroup()
```

```{r, include=T, echo=T}
# calculate haversine distances between from_to points in current table

formulas_d <- formulas %>% 
  rowwise() %>% 
  mutate(dist_haversine = distHaversine(
    c(from_longitude, from_latitude),
    c(to_longitude, to_latitude)) 
      / 1000) %>% 
  ungroup() %>% 
  # make numbers human-readable
  mutate(dist_haversine = round(dist_haversine, digits = 3)) 

head(formulas_d %>% select(lang, text, from_placename, to_placename, dist_haversine))
```

### proportion of dist \> 1000 in lang

```{r, include = T, echo = F}

totals <- formulas_d %>% 
  count(lang) %>% 
  rename(n_total = n)

formulas_d %>% 
  mutate(dist_long = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  count(lang, dist_long) %>% 
  left_join(totals, by = "lang") %>% 
  mutate(perc = round((n/n_total)*100, 1))
```

### distribution

Poet's mind is often flying not that far away?

```{r, include=T, echo=F}
formulas_d %>% 
  #filter(lang == "sl") %>% 
  ggplot(aes(x = lang, y = dist_haversine)) + 
  geom_violin(fill = "lightseagreen") + 
  labs(title = "Distribution of all distances",
       y = "Dist Haversine (km)")
```

```{r, include=T, echo=F}
formulas_d %>% 
  #filter(lang == "sl") %>% 
  ggplot(aes(x = lang, y = dist_haversine, group = lang)) + 
  geom_jitter(alpha = 0.2, width = 0.2) #+
  labs(title = "Distribution of all distances",
       y = "Dist Haversine (km)")
```

What I see here: most of the distances are actually very small; but for some traditions there are quite a portion of longer ones.

The problem is the threshold here: would we want to divide into short/long groups based on any corpus-related number (mean/med dist? country size? which country then), or to set an arbitrary baseline like everything less then 1000km is a small distance, and anything longer – is long?

### Figure 1: dist distribution

```{r}
dist_raw <- formulas_d %>% 
  filter(lang != "it") %>% 
  filter(dist_haversine > 0) %>% 
  ggplot(aes(x = lang, y = dist_haversine, group = lang)) + 
  geom_jitter(alpha = 0.2, width = 0.2, colour = met.brewer("Cassatt2")[10]) + 
  geom_violin(fill = met.brewer("Cassatt2")[4], alpha = 0.7) + 
  labs(y = "Distance (km)",
       x = "Corpus") + 
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

dist_raw
```

```{r}
dist_log <- formulas_d %>% 
  filter(lang != "it") %>% 
  filter(dist_haversine > 0) %>% 
  ggplot(aes(y = factor(lang, levels = c("sl", "ru", "fr", "en", "de", "cs")), 
             x = dist_haversine, group = lang)) + 
  geom_jitter(alpha = 0.2, width = 0.2, colour = met.brewer("Cassatt2")[10]) + 
  geom_violin(fill = met.brewer("Cassatt2")[4], alpha = 0.7) + 
  labs(x = "Distance, km (log scale)",
       y = "Corpus") + 
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14)) + 
  scale_x_log10()

dist_log
```

```{r, eval=F, include=F, echo=F}
plot_grid(dist_raw, NULL, dist_log,
          rel_widths = c(1, 0.05, 1), nrow = 1,
          labels = c("A", "B"), label_size = 18)


ggsave("../plots/Fig_1.png", plot = last_plot(),
       bg = "white", dpi = 300, width = 12, height = 6)
```

### dist summary stats

Calculate mean & median dist for each corpus + 3rd quantile

```{r, include=T, echo=F}
dist_summary <- formulas_d %>% 
  group_by(lang) %>% 
  summarise(dist_mean = mean(dist_haversine),
            dist_median = median(dist_haversine),
            third_quant = quantile(dist_haversine, 0.75)) %>% 
  ungroup()

dist_summary
```

Plot based on the groups long / short distances:

I just wanted to see how these values are distributed, should we divide them in two groups

```{r, include=T, echo=F}
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_longer_median = ifelse(dist_haversine > dist_median, "yes", "no")
         #dist_longer_3qu = ifelse(dist_haversine > third_quant, 1, 0)
         ) %>% 
  select(lang, dist_haversine, dist_longer_median) %>% 
  ungroup() %>% 
  ggplot(aes(x = lang, y = dist_haversine, fill = dist_longer_median)) + 
  geom_boxplot() + 
  labs(title = "Grouping based on: distances longer than median")
```

```{r, include=T, echo=F}
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_longer_3qu = ifelse(dist_haversine > third_quant, "yes", "no")
         ) %>% 
  select(lang, dist_haversine, dist_longer_3qu) %>% 
  ungroup() %>% 
  ggplot(aes(x = lang, y = dist_haversine, fill = dist_longer_3qu)) + 
  geom_boxplot() + 
  labs(title = "Grouping based on: distances longer than 3rd quant limit")
```

Same grouping method for all: dist is short is \< 1000km, long if \>1000km

```{r, include=T, echo=F}
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_longer_1000 = ifelse(dist_haversine > 1000, "yes", "no")
         ) %>% 
  select(lang, dist_haversine, dist_longer_1000) %>% 
  ungroup() %>% 
  ggplot(aes(x = lang, y = dist_haversine, fill = dist_longer_1000)) + 
  geom_boxplot() + 
  labs(title = "Grouped dist distribution: distances longer than 1000km")
```

#### Time: distances based on author birth year

```{r, include = F, echo=F}

# check birth year missing data

totals <- formulas %>% 
  count(lang) %>% 
  rename(total_n = n)

formulas_d %>% 
  select(lang, year_birth, dist_haversine) %>% 
  filter(!is.na(year_birth)) %>% 
  count(lang) %>% 
  left_join(totals, by = "lang")
```

This is how author's birth years are distributed, if we're only looking in our formulas subset

```{r, include=T, echo=F, warning=FALSE}
formulas %>% 
  filter(lang != "it") %>% 
  select(lang, year_birth) %>% 
  ggplot(aes(x = year_birth)) + 
  geom_density(fill = "lightseagreen") + 
  facet_wrap(~lang) + 
  geom_vline(xintercept = seq(1600, 1900, 50), 
             lty = 2, colour = "darkblue") + 
  labs(title = "Formulas found per 50 years period",
       subtitle = "Year of author birth is used")
```

Add time period column & recalculate summary dist metrics: I aggregated based on the half-century periods the author was born. The question here was: do authors born in quite distant times uses different distances when from_to formula appears?

```{r, include=T, echo=F}
dist_t_summary <- formulas_d %>% 
  filter(!is.na(year_birth)) %>% 
  mutate(period_birth = floor(year_birth/50)*50,
         period_end = period_birth + 49,
         period = paste0(period_birth, "—", period_end),
         lang_t = paste0(lang, " ", period)) %>% 
  group_by(lang_t) %>% 
  summarise(dist_mean = mean(dist_haversine),
            dist_median = median(dist_haversine),
            third_quant = quantile(dist_haversine, 0.75),
            number_formulas = n()) %>% 
  ungroup() 

dist_t_summary
```

```{r, include=T, echo=F, warning=F, message=F}
dist_t_summary %>% 
  select(-number_formulas) %>% 
  pivot_longer(!lang_t) %>% 
  left_join(dist_t_summary %>% select(lang_t, number_formulas)) %>% 
  filter(number_formulas > 4) %>% # filter out period if less than [5 formulas found 
  mutate(lang = str_extract(lang_t, "^.."), 
         lang_t = str_remove(lang_t, "^..."),
         lang_t = str_remove(lang_t, "—....$")) %>% 
  
  filter(lang != "it") %>% 
  ggplot(aes(x = lang_t, y = value, group = name, colour = name)) +
  geom_line(aes(linewidth = number_formulas)) + 
  geom_hline(yintercept = 2000, lty = 2, colour = "darkblue") + 
  facet_wrap(~lang, scales = "free_x") + 
  labs(y = "dist in km",
       title = "Temporal distribution of distances in from_to formulas",
       subtitle = "x axis: 50-years chunks based on the author's birth year\n(i.e., 1750: aggregated value for authors born between 1750 and 1799") + 
  theme(legend.position = "bottom")
```

-   overall: we can see than Czech, German, & Slovenian from_to aggregated(!) distances are usually below 2k km; longest from-to are en, fr, and somewhat ru;

-   NB: my 2000 km threshold is absolutely arbitrary; also, I removed periods with less than 5 formulas;

-   Peaks details:

    -   DE: peak period 1700-1749 is only 12 formulas, may be a random result

    -   EN: authors born in 1750-1799: 40 formulas found; for 1800-1849 & 1850-1899 even more data (101 & 143 resp.) –\> these dist are really long on average;

    -   FR: low number of formulas for any poet-birth period EXCEPT for 1800-1849;

    -   RU: most formulas come from authors born in 1850-1899 (69); for 18c-authors there are 15 (1700-1749) and 22 (1750-1799) formulas respectively, so it actually can tell us something – like that these from-to mean/median distances are stable high

### short vs long dist

Previous plot was just aggregating all distances based on 50-year period and calculated means etc. Here I did not do any aggregation, but divided the observations in two groups (\> or \<1000km distance). Then I put everything on roughly the same timeline (no aggregation, just formulas + author's year birth)

```{r, include=T, echo=F, warning=F}
formulas_d %>% 
  filter(lang != "it") %>% 
  mutate(dist_longer_1k = ifelse(dist_haversine > 1000, "yes", "no")) %>% 
  #filter(dist_longer_1k == "no") %>% 
  ggplot(aes(x = year_birth, y = dist_haversine, group = dist_longer_1k)) + 
  geom_point(aes(colour = dist_longer_1k), alpha = 0.5) + 
  facet_wrap(~lang) + 
  geom_smooth(aes(colour = dist_longer_1k, fill = dist_longer_1k)) + 
  geom_hline(yintercept = 5000, lty = 2, colour = "darkblue") + 
  
  theme(legend.position = "bottom") + 
  scale_y_continuous(limits = c(-1, 20000)) + 
  labs(y = "Dist Haversine, km", 
       x = "Author birth year")
```

```{r}
rm(dist_summary, dist_t_summary, loc_freq, dist_log, dist_raw)
```

For longer distances: everyone except EN and FR are usually below the 5k distances even for longer ones? ru is somewhere inbetween.

So I guess the takeouts are:

-   the distribution of from-to distances is very unequal: we actually have more closer distances, when something like "from portugal to china" situations;

-   however the latter is characteristic at least for Eng, Fr, and mb Ru (not surprising but still); for Italy we have too little observations; German corpus looks to me closer to Czech, than to the 'expansion' ones;

-   I think we should consider dividing the distances in groups for sure, otherwise short distances are practically invisible;

-   Question what I'm not sure about is the threshold between short/long distance though, should it be calculated for each country separately, or not, etc. FOR NOW I just used 1000km for all corpora.

# Types

How distances are related to the types of locations in from-to formula?

### Overview

total

```{r, include=T, echo=F}
formulas_d %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  count(type_pair, sort = T) %>% 
  head(20)
```

Add groups of shorter / longer distances (simple 1000 km threshold)

```{r, include=T, echo=F}
total_sl <- formulas_d %>% 
  mutate(dist_type = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  count(dist_type, sort = T)

total_sl_lang <- formulas_d %>% 
  mutate(dist_type = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  count(dist_type, lang, sort = T)

total_sl

total_sl_lang %>% 
  pivot_wider(names_from = dist_type, values_from = n) %>% 
  left_join(totals, by = "lang") %>% 
  mutate(perc_short = round((short/total_n)*100, 1),
         perc_long = round((long/total_n)*100, 1)) %>% 
  select(-total_n)
```

Similar picture to the above: younger literatures are more about closer distances (cs % short – 72%? sl — 87%), while others are more equal to 50/50

### long/short dist ranks

```{r, include=T, echo=F}
formulas_d %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type), 
         dist_type = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  count(type_pair, dist_type, sort = T) %>% 
  group_by(dist_type) %>% 
  #slice_max(order_by = n, n = 10) %>% 
  mutate(rank = row_number()) %>% select(-n) %>% 
  pivot_wider(names_from = dist_type, values_from = rank) %>% 
  arrange(-desc(short)) %>% head(20)

formulas_d %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type), 
         dist_type = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  count(type_pair, dist_type, sort = T) %>% 
  group_by(dist_type) %>% 
  #slice_max(order_by = n, n = 10) %>% 
  mutate(rank = row_number()) %>% select(-n) %>% 
  pivot_wider(names_from = dist_type, values_from = rank) %>% 
  arrange(-desc(long)) %>% head(20)
```

Notes:

-   city –\> city: is a universal thing, can be close, can be distant

-   river –\> river: same but a bit more prevalent for long

-   regions and countries not unexpectedly used for long distances

-   villages & city parts are for short distances

#### by language

```{r, include=T, echo=F}
formulas %>% 
  filter(lang != "it") %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  group_by(lang) %>% 
  count(type_pair, sort = T) %>% 
  filter(n > 1) %>% 
  slice_max(order_by = n, n = 10) %>% 
  ungroup() %>% 
  mutate(type_pair = paste0(type_pair, " (", n, ")")) %>% 
  group_by(lang) %>% 
  mutate(top_list = paste0(type_pair, collapse = " <br> ")) %>% 
  select(-n, -type_pair) %>% 
  ungroup() %>% 
  distinct() %>% 
  #pivot_wider(names_from = lang, values_from = top_list) %>% 
  knitr::kable(escape = F)
```

#### longer vs shorter distances

```{r, include=T, echo=F}
formulas_d %>% 
  filter(lang != "it") %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type), 
         dist_type = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  group_by(lang, dist_type) %>% 
  count(type_pair, sort = T) %>% 
  filter(n > 1) %>% 
  slice_max(order_by = n, n = 5) %>% 
  ungroup() 
```

Also a long table, a summary from me:

-   CZ: mountain-mountain is used both for long and short dist; river-river kind of too, but more for long;
-   DE: same rates for city-city and river-river for both long and short
-   EN: more countries (country-country or city-country); more river-river for shorter dist;
-   FR: long dist for ancient city; shorter: city-city, river-river, city parts and regions
-   RU: long - river-river, city-city; has some short dist for ancient city ? and mountains for short dist too

### ancient cities

```{r, include=T, echo=F}
formulas_d %>% 
  filter(lang != "it") %>% 
  filter(str_detect(from_type, "ancient")) %>% 
  count(lang, from_placename, sort = T)

formulas_d %>% 
  filter(lang != "it") %>% 
  filter(str_detect(to_type, "ancient")) %>% 
  count(lang, to_placename, sort = T)

x <- formulas_d %>% 
  filter(lang != "it") %>% 
  filter(str_detect(from_type, "ancient")) %>% 
  count(lang, from_placename, sort = T) %>% 
  rename(placename = from_placename)

# grouped from and to placenames & filtered >1
formulas_d %>% 
  filter(lang != "it") %>% 
  filter(str_detect(to_type, "ancient")) %>% 
  count(lang, to_placename, sort = T) %>% 
  rename(placename = to_placename) %>% 
  rbind(x) %>% 
  group_by(lang, placename) %>% 
  summarise(total_n = sum(n)) %>% 
  arrange(desc(total_n)) %>% 
  arrange(-desc(lang)) %>% 
  filter(total_n > 1)
```

There are quite a lot of these ancient things, but they are mostly unique (sampled example)

```{r, include=T, echo=F}
formulas_d %>% 
  filter(lang != "it") %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type), 
         dist_type = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  filter(str_detect(type_pair, "ancient")) %>% 
  select(lang, type_pair, text, from_placename, to_placename, dist_type, dist_haversine) %>% 
  sample_n(10)
```

### city parts

```{r, include=T, echo=F}
formulas_d %>% 
  filter(lang != "it") %>% 
  filter(str_detect(from_type, "city part")) %>% 
  count(lang, from_placename, sort = T)

formulas_d %>% 
  filter(lang != "it") %>% 
  filter(str_detect(to_type, "city part")) %>% 
  count(lang, to_placename, sort = T)

# traditions with most city parts present?
x <- formulas_d %>% 
  filter(lang != "it") %>% 
  filter(str_detect(from_type, "city part")) %>% 
  select(lang, from_placename) %>% 
  rename(placename = from_placename)

formulas_d %>% 
  filter(lang != "it") %>% 
  filter(str_detect(to_type, "city part")) %>% 
  select(lang, to_placename) %>% 
  rename(placename = to_placename) %>% 
  rbind(x) %>% 
  group_by(lang) %>% 
  count(placename, sort = T) %>% 
  filter(n > 1) %>% 
  ungroup() %>% 
  arrange(-desc(lang))
```

## Places frequency

### from-to formulas

Just lists of real places, I'm not sure we need to think about this now (not too many observations also)

```{r, include=T, echo=F}
formulas %>% 
  mutate(from_to_pair = paste0(from_placename, " --> ", to_placename)) %>% 
  count(from_to_pair, sort = T) %>% 
  head(20)
```

By language (n \> 1)

```{r, include=T, echo=F}
formulas %>% 
  mutate(from_to_pair = paste0(from_placename, " --> ", to_placename)) %>% 
  group_by(lang) %>% 
  count(from_to_pair, sort = T) %>% 
  slice_max(order_by = n, n = 5, with_ties = F) %>% 
  filter(n > 1) %>% 
  knitr::kable()
```

```{r, eval = F}
formulas_d %>% 
  filter(lang == "fr") %>% 
  filter(from_placename == "Moscow Kremlin" | to_placename == "Moscow Kremlin")
```

### most from / to places

Filter n \> 3 !

```{r, include=T, echo=F}
formulas %>% 
  filter(lang != "it") %>% 
  group_by(lang) %>% 
  count(from_placename, from_type, sort = T) %>% 
  slice_max(order_by = n, n = 5) %>% 
  filter(n > 3)
```

```{r, include=T, echo=F}
formulas %>% 
  filter(lang != "it") %>% 
  group_by(lang) %>% 
  count(to_placename, to_type, sort = T) %>% 
  slice_max(order_by = n, n = 5) %>% 
  filter(n > 3)
```

### most freq places in general

General frequencies (highly dependent on N formulas from a language)

```{r, include=T, echo=F}
x <- formulas %>% 
  select(from_id, from_placename, from_type) %>% 
  rename(id = from_id,
         placename = from_placename,
         type = from_type)

places_freq <- formulas %>% 
  select(to_id, to_placename, to_type) %>% 
  rename(id = to_id, 
         placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  count(id, placename, sort = T) %>% 
  mutate(rank_freq = row_number())

head(places_freq, 20)
```

Types: natural vs political

```{r}
x <- formulas %>% 
  select(from_placename, from_type, lang) %>% 
  rename(placename = from_placename,
         type = from_type)

# look into types
t <- formulas %>% 
  select(to_placename, to_type, lang) %>% 
  rename(placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  pull(type) %>% unique

# create a table for larger groupings:
# whether the entity is natural or political
# if it is water or not
grouping <- tibble(type = t, 
       group = c("p", "n", "n", "p", "n", "p", 
                 "p", "p", "p", "n", "n", "p", 
                 "p", "p", "n", "n", "p", "n",
                 "n", "n", "n", "p"),
       water = c(0, 1, 1, 0, 0, 0, 
                 0, 0, 0, 0, 1, 0, 
                 0, 0, 1, 0, 0, 0, 
                 1, 0, 1, 0))

types_groups <- formulas %>% 
  select(to_placename, to_type, lang) %>% 
  rename(placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  left_join(grouping, by = "type")

totals <- formulas %>% 
  select(to_placename, to_type, lang) %>% 
  rename(placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  count(lang) %>% 
  rename(total = n)

types_groups %>% 
  count(lang, group) %>% 
  left_join(totals, by = "lang") %>% 
  mutate(perc = round((n/total)*100, 1)) %>% 
  select(lang, group, perc) %>% 
  mutate(group = ifelse(group == "n", "Natural", "Political")) %>% 
  pivot_wider(names_from = group, values_from = perc)


types_groups %>% 
  count(lang, water) %>% 
  left_join(totals, by = "lang") %>% 
  mutate(perc = round((n/total)*100, 1)) %>% 
  select(lang, water, perc) %>% 
  mutate(water = ifelse(water == 1, "Water", "Not water")) %>% 
  pivot_wider(names_from = water, values_from = perc)

rm(totals, types_groups, x)
```

Placenames mentioned in all corpora + frequency

```{r, include = T, echo = F}
x <- formulas %>% 
  select(from_placename, from_type, lang) %>% 
  rename(placename = from_placename,
         type = from_type) 

freq_ranks <- formulas %>% 
  select(to_placename, to_type, lang) %>% 
  rename(placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  distinct() %>% 
  count(placename, type, sort = T) %>% 
  rename(n_corpora = n) %>% 
  mutate(rank_corpora = row_number()) %>% 
  left_join(places_freq, by = "placename") %>% 
  rename(freq_abs = n)
  
head(freq_ranks, 20)

# ranks corr
cor(freq_ranks$rank_corpora[1:50], freq_ranks$rank_freq[1:50], method = "kendall")
```

------------------------------------------------------------------------

## Map example: from Malta to Rome

Load pckg for maps

```{r}
library(sf)
library(ggspatial)

library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
```

```{r}
exmpl <- formulas_d %>% 
   filter(str_detect(to_placename, "Rome")) %>% 
   filter(str_detect(author_name, "Whittier")) # specific line in eng

# exmpl <- formulas_d %>% 
#   filter(str_detect(from_placename, "Rhine") & str_detect(to_placename, "Danube"))

exmpl
exmpl$text

exmpl$dist_haversine

exmpl %>% select(from_placename, from_longitude, from_latitude, 
                 to_placename, to_longitude, to_latitude)

```

Basic map

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")

europe <- world[which(world$continent == "Europe"),]

# europe <- world

# from malta to rome
ggplot(europe) +
  geom_sf(fill = "#cbeedb"
          ) +
  coord_sf(xlim = c(5, 20), ylim = c(34,45), expand = FALSE) + 
  
  
  geom_curve(data = exmpl,
             aes(x = from_longitude, y = from_latitude,
                xend = to_longitude, yend = to_latitude),
             linewidth = 3, curvature = 0.2,
             colour = "slategrey", alpha = 0.9) +
  
  # malta label
  geom_text_repel(data = exmpl,
                  aes(x = from_longitude, y = from_latitude, label = from_placename),
                  size = 14, col = "midnightblue", fontface = "bold",
                  vjust = 1.5, 
                  segment.color = "transparent") + 
  
  # rome label
  geom_text_repel(data = exmpl,
                  aes(x = to_longitude, y = to_latitude, label = to_placename),
                  size = 14, col = "violetred4", fontface = "bold",
                  vjust = 1,  box.padding = 0.7,
                  segment.color = "transparent") + 
  
  # malta point
  geom_point(data = exmpl, aes(x = from_longitude, y = from_latitude), 
             size = 7, alpha = 0.9, color = "midnightblue") + 
  
  # rome point
  geom_point(data = exmpl, aes(x = to_longitude, y = to_latitude), 
             size = 7, alpha = 0.9, color = "violetred4") +
  
  
  # annotation_scale(location = "bl", width_hint = 0.4) +
  # annotation_north_arrow(location = "bl", which_north = "true", 
  #       #pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
  #       style = north_arrow_fancy_orienteering)
  
  theme(panel.grid.major = element_line(color = "grey60", 
                                        #linetype = "dashed", 
                                        size = 0.1),
    panel.background = element_rect(fill = "aliceblue"))
```

Line plot

```{r}

exmpl %>% 
  ggplot() + 
  # geom_segment(aes(x = from_longitude, xend = to_longitude,
  #                  y = from_latitude, yend = to_latitude),
  #              color = "slategrey", linewidth = 4
  #              ) +
  geom_point(aes( x = from_longitude, y = from_latitude), 
             color = "midnightblue", size = 10) + 
  geom_point(aes( x = to_longitude, y = to_latitude), 
             color = "violetred4", size = 10) + 
  
  geom_text(aes(x = from_longitude, y = from_latitude),
            label = "Malta", color = "midnightblue",
            size = 10, vjust = -1, hjust = -0.01) + 
  geom_text(aes(x = to_longitude, y = to_latitude),
            label = "Rome", color = "violetred4",
            size = 10, vjust = -1, hjust = -0.01) + 
  
  expand_limits(x = c(10, 16), y = c(35, 45)) + 
  labs(x = "", y = "") + 
  theme(plot.background = element_rect(fill = "transparent",
                                       colour = NA_character_),
        panel.background = element_rect(fill = "transparent", 
                                        colour = NA_character_))
  

ggsave(file = "../plots/chr_lt/2_1.png", plot = last_plot(), bg = "transparent",
       dpi = 300, width = 5, height = 2.6) 
```

Transposed

```{r}
exmpl %>% 
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude) %>% 
  
  ggplot() + 
  
  annotate("point", x = 0, y = 0, size = 200, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, linewidth = 2, colour = "white") + 
  geom_vline(xintercept = 0, lty = 2, linewidth = 2, colour = "white") +

  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "slategrey", linewidth = 4) + 
  
  geom_point(aes(x = x1_0, y = y1_0), 
             colour = "midnightblue", size = 10) + 
  geom_point(aes(x = x2_0, y = y2_0), 
             colour = "violetred4", size = 10) + 
  geom_text(aes(x = x1_0, y = y1_0), 
            colour ="midnightblue", label = "FROM\n(0, 0)",
            vjust = -0.5, size = 10) + 
  geom_text(aes(x = x2_0, y = y2_0), 
            colour ="violetred4", label = "TO",
            vjust = -1, size = 10) + 
  expand_limits(x = c(-4.5, 1.5), y = c(-2, 8)) + 
  labs(x = "", y = "") + 
  theme(plot.background = element_rect(fill = "transparent",
                                       colour = NA_character_),
        panel.background = element_rect(fill = "transparent", 
                                        colour = NA_character_))
  

ggsave(file = "../plots/chr_lt/2_4.png", plot = last_plot(), bg = "transparent",
       dpi = 300, width = 7, height = 4)
  
```

```{r, eval=F, include=F}
# from Rhine to Danube

# ggplot(europe) +
#   geom_sf() +
#   coord_sf(xlim = c(5, 15), ylim = c(43,53), expand = FALSE) + 
#   
#   geom_point(data = exmpl, aes(x = from_longitude, y = from_latitude), 
#              size = 4, alpha = 0.9, color = "steelblue4") + 
#   geom_point(data = exmpl, aes(x = to_longitude, y = to_latitude), 
#              size = 4, alpha = 0.9, color = "seagreen") +
#   
#   geom_curve(data = exmpl,
#              aes(x = from_longitude, y = from_latitude,
#                 xend = to_longitude, yend = to_latitude),
#              linewidth = 1, curvature = 0.2,
#              colour = "seagreen", alpha = 0.7) +
#   geom_text_repel(data = exmpl,
#                   aes(x = from_longitude, y = from_latitude, label = from_placename),
#                   size = 7, col = "steelblue4", fontface = "bold") + 
#   
#   geom_text_repel(data = exmpl,
#                   aes(x = to_longitude, y = to_latitude, label = to_placename),
#                   size = 7, col = "seagreen", fontface = "bold",
#                   vjust = 1,  box.padding = 0.5)
```

## Map: most frequent places

```{r}
loc_fr <- freq_ranks %>% 
  filter(n_corpora > 3) %>% 
  filter(placename == "Baltic Sea")

loc_coord <- formulas %>% 
  filter(from_placename %in% loc_fr$placename | to_placename %in% loc_fr$placename) 

ggplot(europe) +
  geom_sf(fill = "#cbeedb"
          ) +
  coord_sf(xlim = c(-15, 50), ylim = c(35,75), expand = FALSE) + 
  
  geom_curve(data = loc_coord,
             aes(x = from_longitude, y = from_latitude,
                xend = to_longitude, yend = to_latitude),
             linewidth = 0.3, curvature = 0.2,
             colour = "slategrey", alpha = 0.9) + 
  
  geom_point(data = loc_coord, aes(x = from_longitude, y = from_latitude), 
             size = 2, alpha = 0.9, color = "midnightblue") + 
  geom_point(data = loc_coord, aes(x = to_longitude, y = to_latitude), 
             size = 2, alpha = 0.9, color = "violetred4") 
```

```{r}
x <- formulas %>% 
  select(from_id, from_placename, from_latitude, from_longitude, lang) %>% 
  rename(id = from_id,
         placename = from_placename,
         lat = from_latitude,
         long = from_longitude) 

freq_ranks <-formulas %>% 
  select(to_id, to_placename, to_latitude, to_longitude, lang) %>% 
  rename(id = to_id,
         placename = to_placename,
         lat = to_latitude,
         long = to_longitude) %>% 
  rbind(x) %>% 
  distinct() %>% 
  count(id, placename, sort = T) %>% 
  rename(n_corpora = n) %>% 
  filter(n_corpora > 4)

pl_coord <- formulas %>% 
  select(to_id, to_placename, to_latitude, to_longitude, lang) %>% 
  rename(id = to_id,
         placename = to_placename,
         lat = to_latitude,
         long = to_longitude) %>% 
  rbind(x) %>% 
  select(-lang) %>% 
  distinct() %>% 
  filter(placename %in% freq_ranks$placename) 

ggplot(world) +
  geom_sf(#fill = "#cbeedb"
          ) +
  coord_sf(xlim = c(-15, 90), ylim = c(25,75), expand = FALSE) + 
  
  geom_point(data = pl_coord, aes(x = long, y = lat), 
             size = 2, alpha = 0.9, color = "violetred") + 
  geom_text_repel(data = pl_coord, 
                  aes(x = long, y = lat, label = placename),
                  size = 4, color = "midnightblue")

```
