---
title: "02 compass plots"
format: md
editor: visual
---

# Compass plots

```{r, warning=F, include=F}
library(tidyverse)
library(geosphere)

library(cowplot)
library(MetBrewer)
theme_set(theme_minimal())

# maps
library(sf)
library(ggspatial)

library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
```

Load data & calculate distances

```{r, include=T, echo=F}
formulas <- read.csv("../data/formulas_table.csv")
# glimpse(formulas)

formulas_d <- formulas %>% 
  rowwise() %>% 
  mutate(dist_haversine = distHaversine(
    c(from_longitude, from_latitude),
    c(to_longitude, to_latitude)) 
      / 1000) %>% 
  ungroup() %>% 
  # make numbers human-readable
  mutate(dist_haversine = round(dist_haversine, digits = 3),
         dist_longer_1k = ifelse(dist_haversine > 1000, "long", "short")) 

head(formulas_d %>% select(lang, text, from_placename, to_placename, dist_haversine))
```

## calculate coordinates

```{r, eval = F, include = F, echo = F}
# test sample
fs <- formulas %>% sample_n(10)

fs[,9:14]

ggplot() + 
  geom_point(data = fs, aes( x = from_longitude, y = from_latitude)) + 
  geom_point(data = fs, aes( x = to_longitude, y = to_latitude)) + 
  geom_segment(data = fs, aes(x = from_longitude, xend = to_longitude,
                           y = from_latitude, yend = to_latitude))
```

```{r, include=F, eval=F, echo=F}

x1 = fs$from_longitude[1]
y1 = fs$from_latitude[1]

x2 = fs$to_longitude[1]
y2 = fs$to_latitude[1]

x1
y1

x2
y2

x1_0 <- 0
y1_0 <- 0

x2_0 <- x2 - x1
y2_0 <- y2 - y1

x2_0
y2_0
```

```{r, eval=F, include=F, echo=F}
ggplot() + 
  geom_point(data = fs, aes( x = from_longitude, y = from_latitude)) + 
  geom_point(data = fs, aes( x = to_longitude, y = to_latitude)) + 
  geom_segment(data = fs, aes(x = from_longitude, xend = to_longitude,
                           y = from_latitude, yend = to_latitude)) + 
  
  geom_point(aes(x = x1_0, y = y1_0), colour = "red") + 
  geom_point(aes(x = x2_0, y = y2_0), colour = "red") + 
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "red")

```

```{r, eval=F, echo=F, include=F}
fs %>% 
  # calculate new coordinates
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude) %>% 
  glimpse




fs %>% 
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude) %>%
  ggplot() + 
  geom_hline(yintercept = 0, lty = 2, linewidth = 0.7, colour = "lightgray") + 
  geom_vline(xintercept = 0, lty = 2, linewidth = 0.7, colour = "lightgray") + 
  geom_point(aes( x = from_longitude, y = from_latitude)) + 
  geom_point(aes( x = to_longitude, y = to_latitude)) + 
  geom_segment(aes(x = from_longitude, xend = to_longitude, 
                   y = from_latitude, yend = to_latitude)) + 
  
  geom_point(aes(x = x1_0, y = y1_0), colour = "red") + 
  geom_point(aes(x = x2_0, y = y2_0), colour = "red") + 
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "red") 
```

```{r}
formulas_nc <- formulas_d %>% 
  # calculate new coordinates
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude)

glimpse(formulas_nc)
```

## plot — all

```{r, include=T, echo=F}
#p_main <- 
  
formulas_nc %>% 
  filter(lang != "it") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.4, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(5, "pt")),
               alpha = 0.5) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-230, 230)) + 
  scale_y_continuous(limits = c(-230, 230)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 14, face = "bold"),
        # spacing for slides
        panel.spacing = unit(4, "lines"))
  
```

```{r, eval=F, inlude=F, echo=F}
ggsave("../plots/chr_lt/3_1.png", plot = last_plot(), 
       width = 12, height = 10, bg = "white", dpi = 300)
```

## loc types

### river —\> river

```{r, include=T, echo=F}
formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "long") %>% 
  
  filter(from_type == "river" & to_type == "river") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "darkblue",
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-100, 100)) + 
  scale_y_continuous(limits = c(-100, 100)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  labs(title = "River to river directions -- long") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"))
  
```

```{r, include=T, echo=F}
formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "short") %>% 
  
  filter(from_type == "river" & to_type == "river") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "blue",
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-15, 15)) + 
  scale_y_continuous(limits = c(-15, 15)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  labs(title = "River to river directions -- short") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"))
  
```

### city —\> city

```{r, include=T, echo=F}
formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "long") %>% 
  
  filter(from_type == "city" & to_type == "city") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "darkred",
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-150, 150)) + 
  scale_y_continuous(limits = c(-150, 150)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  labs(title = "City to city directions -- long") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"))
  
```

```{r, include=T, echo=F}
formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "short") %>% 
  
  filter(from_type == "city" & to_type == "city") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "darkred",
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-15, 15)) + 
  scale_y_continuous(limits = c(-15, 15)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  labs(title = "City to city directions -- short") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"))
  
```

### country –\> country

```{r, include=T, echo=F}
formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "long") %>% 
  
  filter(from_type == "country" & to_type == "country") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "darkgreen",
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-200, 200)) + 
  scale_y_continuous(limits = c(-200, 200)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  labs(title = "Country to country directions -- long") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"))
  
```

### mountain –\> mountain

```{r, include=T, echo=F}
formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "long") %>% 
  
  filter(from_type == "mountain" & to_type == "mountain") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "purple",
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-200, 200)) + 
  scale_y_continuous(limits = c(-200, 200)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  labs(title = "Mountain to mountain directions -- long") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"))
  
```

### village (any)

```{r, include=T, echo=F}
formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "short") %>% 
  
  filter(from_type == "village" | to_type == "village") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "violet",
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-15, 15)) + 
  scale_y_continuous(limits = c(-15, 15)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  labs(title = "From OR to village directions -- short") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"))
  
```

## compass – 1 lang

Use only one corpus to look into different types of directions

### cz

```{r, include=T, echo=F}
x <- formulas_nc %>% 
  filter(lang == "cs") %>% 
  mutate(loc_type = paste0(from_type, " --> ", to_type)) %>% 
  count(loc_type, sort = T) %>% 
  filter(n > 4) %>% 
  pull(loc_type)
  
x  

formulas_nc %>% 
  filter(lang == "cs") %>% 
  mutate(loc_type = paste0(from_type, " --> ", to_type)) %>% 
  filter(loc_type %in% x) %>% 
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0,
                   colour = loc_type
                   ),
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~loc_type, ncol = 3) + 
  scale_x_continuous(limits = c(-130, 130)) + 
  scale_y_continuous(limits = c(-130, 130)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  # labs(title = "River to river") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold")) + 
  labs(title = "Only CZ corpus",
       subtitle = "(loc_type with n > 4)")
  
```

NB here longest distances are not only river-river, but mountains

### en

```{r, include=T, echo=F}
x <- formulas_nc %>% 
  filter(lang == "en") %>% 
  mutate(loc_type = paste0(from_type, " --> ", to_type)) %>% 
  count(loc_type, sort = T) %>% 
  filter(n > 4) %>% 
  pull(loc_type)
  
x  

formulas_nc %>% 
  filter(lang == "en") %>% 
  mutate(loc_type = paste0(from_type, " --> ", to_type)) %>% 
  filter(loc_type %in% x) %>% 
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0,
                   colour = loc_type
                   ),
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~loc_type, ncol = 3) + 
  scale_x_continuous(limits = c(-230, 230)) + 
  scale_y_continuous(limits = c(-230, 230)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  # labs(title = "River to river") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold")) + 
  labs(title = "Only EN corpus",
       subtitle = "(loc_type with n > 4)")
  
```

### fr

```{r, include=T, echo=F}
x <- formulas_nc %>% 
  filter(lang == "fr") %>% 
  mutate(loc_type = paste0(from_type, " --> ", to_type)) %>% 
  count(loc_type, sort = T) %>% 
  filter(n > 3) %>% 
  pull(loc_type)
  
x  

formulas_nc %>% 
  filter(lang == "fr") %>% 
  mutate(loc_type = paste0(from_type, " --> ", to_type)) %>% 
  filter(loc_type %in% x) %>% 
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0,
                   colour = loc_type
                   ),
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~loc_type, ncol = 3) + 
  scale_x_continuous(limits = c(-150, 150)) + 
  scale_y_continuous(limits = c(-150, 150)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  # labs(title = "River to river") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold")) + 
  labs(title = "Only FR corpus",
       subtitle = "(loc_type with n > 3)")
  
```

### ru

```{r, include=T, echo=F}
x <- formulas_nc %>% 
  filter(lang == "ru") %>% 
  mutate(loc_type = paste0(from_type, " --> ", to_type)) %>% 
  count(loc_type, sort = T) %>% 
  filter(n > 3) %>% 
  pull(loc_type)
  
x  

formulas_nc %>% 
  filter(lang == "ru") %>% 
  mutate(loc_type = paste0(from_type, " --> ", to_type)) %>% 
  filter(loc_type %in% x) %>% 
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0,
                   colour = loc_type
                   ),
               arrow = arrow(length = unit(4, "pt"))) + 
  facet_wrap(~loc_type, ncol = 3) + 
  scale_x_continuous(limits = c(-230, 230)) + 
  scale_y_continuous(limits = c(-230, 230)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  # labs(title = "River to river") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold")) + 
  labs(title = "Only ru corpus",
       subtitle = "(loc_type with n > 3)")
  
```

## add short/long distances types

two different plots for long & for short distances

### short dist closer look

```{r, include=T, echo=F}
p_short <- formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "short") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 70, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[4], 
           colour = met.brewer("Cassatt1")[4]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(4, "pt")),
               alpha = 0.5) + 
  # facet_wrap(~lang, ncol = 1) + 
  scale_x_continuous(limits = c(-15, 15)) + 
  scale_y_continuous(limits = c(-15, 15)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 14, face = "bold")) + 
  labs(title = "Short distances (<1000 km)")
  
p_short + facet_wrap(~lang, ncol = 3)

p1 <- p_short + facet_wrap(~lang, ncol = 1)
```

```{r, eval=F, inlude=F, echo=F}
ggsave("../plots/01_compass_dist_short.png", plot = last_plot(), 
       width = 10, height = 8, bg = "white", dpi = 300)
```

### longer dist

```{r, include=T, echo=F}
p_long <- formulas_nc %>% 
  filter(lang != "it") %>% 
  filter(dist_longer_1k == "long") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.4, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(4, "pt")), 
               alpha = 0.5) + 

  scale_x_continuous(limits = c(-230, 230)) + 
  scale_y_continuous(limits = c(-230, 230)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 14, face = "bold")) + 
  labs(title = "Long distances (>1000 km)")


p_long + facet_wrap(~lang, ncol = 3)  

p2 <-  p_long + facet_wrap(~lang, ncol = 1)
```

```{r, eval=F, inlude=F, echo=F}
ggsave("../plots/02_compass_dist_long.png", plot = last_plot(), 
       width = 15, height = 11, bg = "white", dpi = 300)
```

Merge two plots in one

```{r, eval=F, include=F, echo=F}
plot_grid(p1, p2)

ggsave("../plots/Fig_3.png", plot = last_plot(), 
       bg = "white", dpi = 300, width = 8, height = 20)
```

### Directions in time

Temporal facets for each corpus: groupings by author's birth year; 50-years spans

```{r}

# fucntion for easier running

time_plot <- function(corpus, lim_x1, lim_x2, lim_y1, lim_y2, point_size) { 
  
  formulas_nc %>% 
  filter(!is.na(year_birth)) %>% 
  mutate(period_birth = floor(year_birth/50)*50,
         period_end = period_birth + 49,
         period = paste0(period_birth, "—", period_end)) %>% 
  filter(lang == corpus) %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = point_size, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  facet_wrap(~period, ncol=6) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(4, "pt"))) + 
  scale_x_continuous(limits = c(lim_x1, lim_x2)) + 
  scale_y_continuous(limits = c(lim_y1, lim_y2)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 12, face = "bold"))
  }
```

Czech corpus:

```{r, include=T, echo=F}
time_plot(corpus = "cs", 
          lim_x1 = -150, lim_x2 = 150, 
          lim_y1 = -150, lim_y2 = 150,
          point_size = 80) + labs(title = "Czech")
```

German corpus: there are mostly small distances which are hardly visible here

```{r, include=T, echo=F}
time_plot(corpus = "de", 
          lim_x1 = -140, lim_x2 = 140, 
          lim_y1 = -140, lim_y2 = 140,
          point_size = 80) + labs(title = "DE")
```

English: interesting expansion to north and to south, huh? and later to the north

```{r, include=T, echo=F}
time_plot(corpus = "en", 
          lim_x1 = -230, lim_x2 = 230, 
          lim_y1 = -230, lim_y2 = 230,
          point_size = 80) + labs(title = "EN")
```

FR: authors born in the first half of the 19th-c. are colonialists

```{r, include=T, echo=F}
time_plot(corpus = "fr", 
          lim_x1 = -200, lim_x2 = 200, 
          lim_y1 = -200, lim_y2 = 200,
          point_size = 80) + labs(title = "FR")
```

IT: the only case of reducing the scope of from-to

```{r, include=T, echo=F}
time_plot(corpus = "it", 
          lim_x1 = -150, lim_x2 = 150, 
          lim_y1 = -150, lim_y2 = 150,
          point_size = 80) + labs(title = "IT")
```

RU: horizontal expansion + caucasus?

```{r, include=T, echo=F}
time_plot(corpus = "ru", 
          lim_x1 = -190, lim_x2 = 190, 
          lim_y1 = -190, lim_y2 = 190,
          point_size = 80) + labs(title = "RU")
```

SL: too small (3 longest dist are removed on this plot)

```{r, include=T, echo=F}
time_plot(corpus = "sl", 
          lim_x1 = -15, lim_x2 = 15, 
          lim_y1 = -15, lim_y2 = 15,
          point_size = 80) + labs(title = "SL")
```

# most freq places

```{r}
x <- formulas_nc %>% 
  select(from_placename, from_type, lang) %>% 
  rename(placename = from_placename,
         type = from_type) 

freq_ranks <- formulas_nc %>% 
  select(to_placename, to_type, lang) %>% 
  rename(placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  distinct() %>% 
  count(placename, type, sort = T) %>% 
  rename(n_corpora = n) %>% 
  filter(n_corpora > 3) # %>% 
  # filter(placename %in% c(
  #   "Baltic Sea", "Tagus River", "Paris", "Rome", "Athens", 
  #   "Nile", "Euphrates", "Babylon", "Bethlehem", 
  #   "Dardanelles", "Don", "Moscow"
  # )) 

freq_ranks


loc_coord <- formulas_nc %>% 
  filter(from_placename %in% freq_ranks$placename) 

freq_coord <- loc_coord %>% 
  left_join(freq_ranks %>% 
              mutate(grouping = ifelse(placename %in% c("Paris", "Rome", "Vienna"),
                                                    "center", "border")) %>% 
              select(placename, grouping) %>% 
              rename(from_placename = placename), 
            by = "from_placename") # %>% 
  # left_join(freq_ranks %>% select(placename, grouping) %>% 
  #             rename(to_placename = placename,
  #                    grouping_t = grouping)) %>% 
  # mutate(grouping = ifelse(is.na(grouping), grouping_t, grouping)) %>% 
  # select(-grouping_t)

freq_coord
```

## map

```{r}
# load map
world <- ne_countries(scale = "medium", returnclass = "sf")

# europe <- world[which(world$continent == "Europe"),]


x <- freq_coord %>% 
  select(from_id, from_placename, from_latitude, from_longitude, lang) %>% 
  rename(id = from_id,
         placename = from_placename,
         lat = from_latitude,
         long = from_longitude) 


pl_coord <- freq_coord %>% 
  select(to_id, to_placename, to_latitude, to_longitude, lang) %>% 
  rename(id = to_id,
         placename = to_placename,
         lat = to_latitude,
         long = to_longitude) %>% 
  rbind(x) %>% 
  select(-lang) %>% 
  distinct() %>% 
  filter(placename %in% freq_ranks$placename) 

ggplot(world) +
  geom_sf(#fill = "#cbeedb"
          ) +
  coord_sf(xlim = c(-15, 80), ylim = c(27,62), expand = FALSE) + 
  
  geom_point(data = pl_coord, aes(x = long, y = lat), 
             size = 2, alpha = 0.9, color = "violetred") + 
  geom_text_repel(data = pl_coord, 
                  aes(x = long, y = lat, label = placename),
                  size = 4, color = "midnightblue")

```

Volga instead of Moscow

check compasses & change colours on compasses (borders / centers wrong direction) – leave only Rome as center?

prob make compasses for all loc in 4+ corp & try

```{r}
ggplot(world) +
  geom_sf(fill = "cornsilk1",
          color = "cornsilk4",
          linewidth = 1
          ) +
  coord_sf(xlim = c(-15, 68), ylim = c(27,62), expand = FALSE) + 
  
  geom_point(data = pl_coord, aes(x = long, y = lat),
             size = 1, alpha = 0.9, color = "violetred4") + 
  theme(panel.grid.major = element_line(color = "grey60", 
                                        #linetype = "dashed", 
                                        linewidth = 0.1), 
        panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(color = "grey60"))  
  
  # geom_text_repel(data = pl_coord, 
  #                 aes(x = long, y = lat, label = placename),
  #                 size = 4, color = "midnightblue")
  #annotation_scale(location = "bl", width_hint = 0.2) # +
  # annotation_north_arrow(location = "bl", which_north = "true", 
  #        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
  #        width = unit(1, "cm"), height = unit(1, "cm"), 
  #        style = north_arrow_fancy_orienteering)

ggsave(file = "../plots/chr_lt/3_0.png", plot = last_plot(), 
       bg = "white",
       dpi = 300, width = 20, height = 10)
```

## compasses

Only compasses for the places appeared in all 6 or 5 corpora

```{r}
glimpse(freq_coord)
```

```{r}
freq_coord %>% 
  mutate(base_place = ifelse(from_placename %in% freq_ranks$placename, 
                             from_placename, to_placename)) %>% 
  count(base_place, sort = T)
```

```{r}

fixed_length <- 100

freq_coord %>% 
  # filter only from places
  filter(lang != "it") %>% 
  mutate(base_place = ifelse(from_placename %in% freq_ranks$placename, 
                             from_placename, "")) %>% 
  filter(base_place != "") %>% 
  
  # calculate vector length (hypotenuse) & normalize it
  mutate(
    # hypotenuse
    vec_len = sqrt(x2_0^2 + y2_0^2),
    
    # norm
    x2_norm = (x2_0 / vec_len) * fixed_length,
    y2_norm = (y2_0 / vec_len) * fixed_length
         ) %>% 
  
  # plot
  ggplot() + 
  
  annotate("point", x = 0, y = 0, size = 120,
           alpha = 0.4,
           fill = met.brewer("Cassatt1")[5],
           colour = met.brewer("Cassatt1")[5]) +

  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_norm,
                   y = y1_0, yend = y2_norm,
                   color = grouping), 
               #colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(7, "pt")),
               size = 2,
               alpha = 1) + 
  facet_wrap(~base_place, ncol = 3) + 
  scale_color_manual(values = c(met.brewer("Cassatt2")[9], 
                                met.brewer("Cassatt1")[1])) +
  scale_x_continuous(limits = c(-120, 120)) + 
  scale_y_continuous(limits = c(-120, 120)) + 
  
  annotate("point", x = 0, y = 0, size = 7,
           fill = "violetred4", colour = "violetred4") + 
  
  theme(# plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 14, face = "bold"), 
        
        plot.background = element_rect(fill = "transparent",
                                       colour = NA_character_),
        panel.background = element_rect(fill = "transparent", 
                                        colour = NA_character_))
  
ggsave(file = "../plots/chr_lt/3_1.png", plot = last_plot(), 
       bg = "transparent",
       dpi = 300, width = 15, height = 39)
```
