---
title: "03_directions_modelling"
format: html
editor: visual
---

## Directions modelling

The notebook takes centralized geo-data from the from—to formulas and transform it to summarize major directions and their dependence on location types

```{r}
library(tidyverse)
library(geosphere) # for Haversine dist calculation

library(cowplot)
library(MetBrewer)
theme_set(theme_minimal())
```

```{r, include=T, echo=F}
formulas <- read.csv("../data/formulas_table.csv")
glimpse(formulas)

formulas_d <- formulas %>% 
  filter(lang != "it") %>% 
  rowwise() %>% 
  mutate(dist_haversine = distHaversine(
    c(from_longitude, from_latitude),
    c(to_longitude, to_latitude)) 
      / 1000) %>% 
  ungroup() %>% 
  # make numbers human-readable
  mutate(dist_haversine = round(dist_haversine, digits = 3),
         dist_longer_1k = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  distinct() %>%  # remove duplicated from-to from the same poem
  
  # calculate new coordinates
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude) 

head(formulas_d %>% select(lang, text, from_placename, to_placename, dist_haversine))

rm(formulas)
```

## types edit

Create larger groups for location types

```{r}
unique(c(formulas_d$from_type, formulas_d$to_type))

types_groups <- tibble(type_original = unique(c(formulas_d$from_type, formulas_d$to_type)),
       type_grouped = c("water", "other", "land", 
                        "city", "water", "mountain", 
                        "country", "water", "settlement", 
                        "city", "settlement", "city", 
                        "water", "land", "land", 
                        "land", "mountain", "water", 
                        "settlement", "water", "mountain", 
                        "settlement"), 
       type_major = c("natural", "natural", "natural", 
                      "political", "natural", "natural", 
                      "political", "natural", "political", 
                      "political", "political", "political", 
                      "natural", "natural", "political", 
                      "natural", "natural", "natural", 
                      "political", "natural", "natural", 
                      "political"))

# attach to the main table for from- and to-locs
formulas_d <- formulas_d %>% 
  left_join(types_groups %>% 
              rename(
                from_type = type_original,
                from_type_group = type_grouped,
                from_type_major = type_major),
            by = "from_type") %>% 
  left_join(types_groups %>% 
              rename(
                to_type = type_original,
                to_type_group = type_grouped,
                to_type_major = type_major),
            by = "to_type") 
```

```{r}
glimpse(formulas_d)
```

```{r}
formulas_d %>% count(from_type_group)
formulas_d %>% count(to_type_group)

formulas_d %>% count(from_type_major)
formulas_d %>% count(to_type_major)
```

## Angle calculation

The idea is to calculate the angle between the clock handle to 12 o'clock (north) and the from-to vector. The default calculation start counting angle from the x-axis (east) & counter-clockwise. So there are some additional steps to calculate this as if we had a clock, that NORTH=0°

### Sanity check

```{r}
t <- formulas_d %>% 
  mutate(angle_east_deg = atan2(y2_0, x2_0) * 180 / pi,
         angle_north_deg = (90 - angle_east_deg + 360) %% 360) %>% 
  filter(dist_longer_1k == "long") %>% 
  sample_n(6)

t

t %>% 
  ggplot() + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "lightblue", 
             linewidth = .5) + 
  geom_vline(xintercept = 0, lty = 2, colour = "lightblue", 
             linewidth = .5) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0)) + 
  geom_point(aes(x = rep(0, 6), y = rep(0, 6))) + 
  
  geom_text(aes(x = x2_0+10, y = y2_0+10, label = round(angle_north_deg)),
            colour = "lightpink") + 
  facet_wrap(~doc_key, scales = "fixed") + 

  scale_x_continuous(limits = c(-90, 90)) +
  scale_y_continuous(limits = c(-90, 90)) +
  coord_equal()
```

```{r}
t
```

```{r}
(90 - (atan2(13, -18) * 180 / pi) + 360) %% 360
```

Calculate angles

```{r}
formulas_d <- formulas_d %>% 
  mutate(angle_east_deg = atan2(y2_0, x2_0) * 180 / pi,
         angle_north_deg = (90 - angle_east_deg + 360) %% 360) 
```

## Compass labels

Use `case_when`\`for separate into N, NE, E, SE, S, SW, W, NW groups, 45 degrees each.

```{r}
formulas_d <- formulas_d %>% 
  mutate(
    compass = case_when(
      angle_north_deg < 22.5 | angle_north_deg >= 337.5 ~ "N",
      angle_north_deg < 67.5  ~ "NE",
      angle_north_deg < 112.5 ~ "E",
      angle_north_deg < 157.5 ~ "SE",
      angle_north_deg < 202.5 ~ "S",
      angle_north_deg < 247.5 ~ "SW",
      angle_north_deg < 292.5 ~ "W",
      TRUE                    ~ "NW"
    )
  )
```

### compass viz

```{r}
# factor level order
formulas_d <- formulas_d %>% 
  mutate(
    compass = factor(
      compass,
      levels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW")
    )
  )
```

```{r}
dist_mean <- formulas_d %>% 
  group_by(lang) %>% 
  summarise(dist_mean = mean(dist_haversine),
            .groups = "drop")

totals <- formulas_d %>% 
  left_join(dist_mean, by = "lang") %>% 
  mutate(dist_longer_mean = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  count(lang, dist_longer_mean) %>% 
  mutate(lang_d = paste0(lang, "_", dist_longer_mean)) %>% 
  rename(total_f = n)

totals

x <- formulas_d %>% 
  left_join(dist_mean, by = "lang") %>% 
  mutate(dist_longer_mean = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(lang_d = paste0(lang, "_", dist_longer_mean)) %>% 
  group_by(lang_d) %>% 
  count(compass) %>% 
  left_join(totals, by = "lang_d") %>% 
  mutate(perc = (n / total_f)) %>% 
  ungroup() %>% 
  mutate(value = 1)  
  
x
```

```{r}
long_dist_pizza <- x %>% 
  filter(dist_longer_mean == "long" & lang != "it") %>% 
  ggplot(aes(x = compass, y = value)) + 
    geom_col(width = 1,
           position = "fill", colour = "white",
           fill = met.brewer("Cassatt1")[8],
           aes(alpha = perc)) + 
    facet_wrap(~lang) + 
    labs(alpha = "%") + 
    geom_hline(yintercept = 1.0, colour = met.brewer("Cassatt1")[8], alpha = 0.5) + 
    theme(
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.text.y = element_blank(),
      strip.text = element_text(face = "bold", size = 20),
      #legend.position = "bottom",
      legend.text = element_text(size = 14),
      legend.title = element_text(size = 14),
      axis.text.x = element_text(face = "bold", size = 12)
      ) + 
    coord_polar(start = -pi/8) 

long_dist_pizza
```

```{r}
short_dist_pizza <- x %>% 
  filter(dist_longer_mean == "short" & lang != "it") %>% 
  ggplot(aes(x = compass, y = value)) + 
  geom_col(width = 1,
           position = "fill", colour = "white",
           fill = met.brewer("Cassatt2")[3],
           aes(alpha = perc)) + 
  facet_wrap(~lang) + 
  labs(alpha = "%") +
  geom_hline(yintercept = 1.0, colour = met.brewer("Cassatt2")[3], alpha = 0.5) + 
  theme(
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.text.y = element_blank(),
      strip.text = element_text(face = "bold", size = 20),
      #legend.position = "bottom",
      legend.text = element_text(size = 14),
      legend.title = element_text(size = 14),
      axis.text.x = element_text(face = "bold", size = 12)
      ) + 
  coord_polar(start = -pi/8) 

short_dist_pizza
```

## Model

This is the resulting data that can be used for the model:

`angle_north_deg` is the angle used for the compass, `compass` column is for the labels.

Columns `from/to_type_group` and `from/to_type_major` contain the labels for groups (city, water, etc.) and binary classes as natural/political

```{r}
glimpse(formulas_d)
```

```{r}
formulas_d %>% count(from_type_group)
```

```{r}
formulas_d %>% count(from_type_major)
```

Distribution of angle instances

```{r}
formulas_d %>% 
  ggplot(aes(x = round(angle_north_deg,0))) + geom_bar() + 
  facet_wrap(~lang)
```

Same but labels + types (for Czech)

```{r}
formulas_d %>% 
  filter(lang == "cs") %>% 
  group_by(dist_longer_1k) %>% 
  count(compass) %>% 
  ungroup() %>% 
  ggplot(aes(x = compass, y = n, fill = compass)) + 
  geom_col() + 
  facet_wrap(~dist_longer_1k, ncol = 1) + 
  scale_fill_manual(values = c(rev(met.brewer("Cassatt1"))[1:2],
                               rev(met.brewer("Cassatt1"))[5:7],
                               (met.brewer("Cassatt1"))[4:8]))
```
