---
title: "04_borders_model"
format: html
---

## Direction delta's

**!! NB !!** will need `formulas_d` dataset from `03_directions_modelling.qmd`.


The idea is to quantify how "structured" are different types of location   
(are cities really centers? are bodies of water really borders?)

Step 1. Take a location (e.g. Rome). Calculate all *differences* between outgoing directions
Step 2. Measure "spikiness"/"entropy" of a resulting distribution. This is our estimand E.
Step 3. Each location has E: we want to ask what influences the structure: does "borderness" depend on the type of the location.
Step 4. Fit a model: "E ~ type * log(n_occurences) + n_languages"

n_occurences controls for obvious increase in structure with decrease of observations.

```{r}
library(ineq)
from_select <- formulas %>% count(from_id,sort=T) %>% filter(n>2)
to_select <- formulas %>% count(from_id,sort=T) %>% filter(n>2)

unique(c(formulas_d$from_type, formulas_d$to_type))

types_groups <- tibble(type_original = unique(c(formulas_d$from_type, formulas_d$to_type)),
       type_grouped = c("water", "other", "land", 
                        "city", "water", "mountain", 
                        "country", "water", "settlement", 
                        "city", "settlement", "city", 
                        "water", "land", "land", 
                        "land", "mountain", "water", 
                        "settlement", "water", "mountain", 
                        "settlement"), 
       type_major = c("natural", "natural", "natural", 
                      "political", "natural", "natural", 
                      "political", "natural", "political", 
                      "political", "political", "political", 
                      "natural", "natural", "political", 
                      "natural", "natural", "natural", 
                      "political", "natural", "natural", 
                      "political"))

# attach to the main table for from- and to-locs
formulas_d <- formulas_d %>% 
  left_join(types_groups %>% 
              rename(
                from_type = type_original,
                from_type_group = type_grouped,
                from_type_major = type_major),
            by = "from_type") %>% 
  left_join(types_groups %>% 
              rename(
                to_type = type_original,
                to_type_group = type_grouped,
                to_type_major = type_major),
            by = "to_type") 


# prepare a list of DF holding all "FROM A" directions
df_list <- formulas_d %>%
  filter(from_id %in% from_select$from_id) %>% mutate(diff_entropy = 0,diff_gini=0) %>%  group_by(from_id) %>% group_split()

df_list_to <- formulas_d %>%
  filter(to_id %in% to_select$to_id) %>% mutate(diff_entropy = 0,diff_gini=0) %>%  group_by(to_id) %>% group_split()

for(i in 1:length(df_list)) {
  
  df <- df_list[[i]]
  ## circular absolute distance between angles
  circ_dist <- abs(
    (
      (outer(df$angle_east_deg, df$angle_east_deg, "-") + 180) %% 360) - 180
)
  ## only lower
  distances <- circ_dist[lower.tri(circ_dist)]
  
  ## bin values 
  bins <- seq(0, 180, by = 5) 
  hist_vals <- hist(distances, breaks = bins, plot = FALSE)
  
  ## probabilities of a differences over binned 0,180 interval
  p <- hist_vals$counts / sum(hist_vals$counts)
  
  ## gini and entropy
  df <- df %>% mutate(diff_entropy = -sum(p[p > 0] * log(p[p > 0])),
                      diff_gini = ineq(p, type="Gini"))
  
  df_list[[i]] <- df
  

}


formulas_ineq <- df_list %>%
  bind_rows() %>% 
  group_by(from_id) %>% 
  mutate(n = max(row_number()), # number of times a place appears
         n_langs = length(unique(lang))) %>%  # number of languages referencing the place
  ungroup() %>% 
  select(lang, n_langs,n,starts_with("from_"), starts_with("diff_")) %>%
  filter(!duplicated(from_id)) # table for the model

cor(formulas_ineq$diff_entropy,formulas_ineq$diff_gini)


formulas_to_ineq <- df_list_to %>%
  bind_rows() %>% 
  group_by(to_id) %>% 
  mutate(n = max(row_number()), # number of times a place appears
         n_langs = length(unique(lang))) %>%  # number of languages referencing the place
  ungroup() %>% 
  select(lang, n_langs,n,starts_with("to_"), starts_with("diff_")) %>%
  filter(!duplicated(to_id)) # table for the model


```

```{r}

intersection <- intersect(formulas_ineq$from_placename,formulas_to_ineq$to_placename)
union1 <- union(formulas_ineq$from_placename,formulas_to_ineq$to_placename)

length(intersection)
length(union1)

top_from <- formulas_ineq %>% filter(n>=8)
top_to <- formulas_to_ineq %>% filter(n>=6)

## changes in structure?
top_from %>% 
  left_join(top_to,by = c("from_id" = "to_id")) %>%
  filter(from_placename==to_placename) %>% 
  mutate(delta = diff_entropy.x - diff_entropy.y) %>%
  select(from_placename,delta) %>%
  arrange(-abs(delta)) 
```


### Fitting the model

```{r}
library(brms)

## regularizing priors
priors <- c(
  prior(normal(0, 0.5), class = "b"),
  prior(normal(1, 0.5), class = "Intercept")
)

## prior predictive check (just make sure it's reasonable enough)
fit_prior <- brm(
  diff_entropy ~ from_type_major*log(n) + n_langs,
  data = formulas_ineq,
  family = gaussian(),
  prior = priors,
  sample_prior = "only"
)

pp_check(fit_prior)


m0m2 <- brm(
  diff_entropy ~ from_type_major*log(n) + n_langs,
  data = formulas_ineq,
  family = gaussian(),
  chains = 4,
  cores = 4,
  prior = priors,
  iter = 4000,
  save_model = "m0m2",
)

saveRDS(m0m2,file="../models/entropy_major_type.rds")


m0m2_to <- brm(
  diff_entropy ~ to_type_major*log(n) + n_langs,
  data = formulas_to_ineq,
  family = gaussian(),
  chains = 4,
  cores = 4,
  prior = priors,
  iter = 4000,
  save_model = "m0m2_to",
)

saveRDS(m0m2_to,file="../models/entropy_to_major_type.rds")



```

```{r}
m0m2 <- readRDS("../models/entropy_major_type.rds")
m0m2_to <- readRDS("../models/entropy_to_major_type.rds")
## from
pp_check(m0m2)
conditional_effects(m0m2)

summary(m0m2)

## to
pp_check(m0m2_to)
conditional_effects(m0m2_to)

summary(m0m2_to)

```

### Plots

Redrawing conditional effects here:

```{r}
ce <- conditional_effects(m0m2)
ce_df <- as_tibble(ce[["n:from_type_major"]])



p3 <- ggplot(ce_df,aes(x = n, y = estimate__, color = from_type_major, fill = from_type_major)) +
  geom_line() +
  geom_ribbon(
    aes(ymin = lower__, ymax = upper__),
    alpha = 0.2,
    color = NA
  ) +
  theme_minimal(base_size=12) +
  labs(
    x = "Number of occurrences of a location",
    y = "Predicted entropy"
  ) +
  geom_vline(aes(xintercept=10),linetype=2,color="grey40") + 
  scale_color_manual(values=met.brewer("Cassatt1")[c(2,8)]) +
  scale_fill_manual(values=met.brewer("Cassatt1")[c(2,8)]) + labs(color="Location type",fill="Location type",title="Interaction between type and number of locations")
``` 


```{r}
# Get all levels of type
types <- unique(formulas_ineq$from_type_major)

# Build new data frame
newdata <- tibble(
  from_type_major = types,
  n = 10,
  n_langs = 3)


posterior_draws <- posterior_predict(m0m2, newdata = newdata)
posterior_means <- posterior_epred(m0m2, newdata = newdata)


summary_df <- as_tibble(posterior_means)
colnames(summary_df) <- types


# average diff in means ~0.43 entropy points
mean(unlist(summary_df[,1] - summary_df[,2]))

pp_diff <- table((summary_df[,1] - summary_df[,2]) > 0)
# 96% of the time "political" predictions have larger mean
pp_diff[2]/sum(pp_diff)

means1 <- summary_df %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "pred") %>% 
  group_by(type) %>% summarize(m1=mean(pred))

p1 <- summary_df %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "pred") %>%
  ggplot(aes(pred,reorder(type,-pred),color=type,fill=type)) +
  
  geom_density_ridges(size=0.5,alpha=0.3,scale=1) + 
  geom_vline(data=means1, aes(xintercept=m1,color=type),linetype=2) +
  theme_minimal(base_size = 12) + 
  scale_color_manual(values=met.brewer("Cassatt1")[c(2,8)]) +
  scale_fill_manual(values=met.brewer("Cassatt1")[c(2,8)]) + 
  guides(color="none",fill="none") +
  labs(x="",y=NULL,title="FROM A locations")
```
"To" distances

```{r}
# Get all levels of type
types <- unique(formulas_to_ineq$to_type_major)

# Build new data frame
newdata <- tibble(
  to_type_major = types,
  n = 10,
  n_langs = 3)


posterior_draws <- posterior_predict(m0m2_to, newdata = newdata)
posterior_means <- posterior_epred(m0m2_to, newdata = newdata)


summary_df <- as_tibble(posterior_means)
colnames(summary_df) <- types


# average diff in means ~0.3 entropy points
mean(unlist(summary_df[,1] - summary_df[,2]))

pp_diff <- table((summary_df[,1] - summary_df[,2]) > 0)
# 95% of the time "political" predictions have larger mean
pp_diff[2]/sum(pp_diff)


means1 <- summary_df %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "pred") %>% 
  group_by(type) %>% summarize(m1=mean(pred))

p2 <- summary_df %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "pred") %>%
  ggplot(aes(pred,reorder(type,-pred),color=type,fill=type)) +
  
  geom_density_ridges(size=0.5,alpha=0.3,scale=1) + 
  geom_vline(data=means1, aes(xintercept=m1,color=type),linetype=2) +
  theme_minimal(base_size = 12) + 
  labs(x="Predicted marginal entropy",y=NULL,title="TO B locations") +
  scale_color_manual(values=met.brewer("Cassatt1")[c(2,8)]) +
  scale_fill_manual(values=met.brewer("Cassatt1")[c(2,8)]) + labs(color="Location type",fill="Location type") + guides(color="none",fill="none")


library(patchwork)

(p1 / p2) | p3

ggsave("../plots/Fig_model.png",height = 6,width = 10)
```

## Rivers: robustness check

Here we refit model based on differently coded river coordinates (wikidata api, mouth, delta, "midpoint").  

1. Each river can be represented by either mouth, delta, or midpoint coordiantes
2. We calculate angles, clock hands, entropy and Gini based on randomly sampled coord for rivers (in "to" and "from" positions)
3. Update the model based on new data
4. Repeat; check posterior effect.

### Wrangling and Recalculating data


```{r}
formulas <- read_csv("../data/formulas_table.csv")
## river coords
rivers <- read_csv("../data/working_files/rivers_with_coords.csv") %>% 
  select(id,source_lat,source_lon, mouth_lat, mouth_lon, mid_lat,mid_lon)

m0m2 <- readRDS("../models/entropy_major_type.rds")
m0m2_to <- readRDS("../models/entropy_to_major_type.rds")

from_select <- formulas %>% count(from_id,sort=T) %>% filter(n>2)
to_select <- formulas %>% count(from_id,sort=T) %>% filter(n>2)


types_groups <- tibble(type_original = unique(c(formulas$from_type, formulas$to_type)),
       type_grouped = c("water", "other", "land", 
                        "city", "water", "mountain", 
                        "country", "water", "settlement", 
                        "city", "settlement", "city", 
                        "water", "land", "land", 
                        "land", "mountain", "water", 
                        "settlement", "water", "mountain", 
                        "settlement"), 
       type_major = c("natural", "natural", "natural", 
                      "political", "natural", "natural", 
                      "political", "natural", "political", 
                      "political", "political", "political", 
                      "natural", "natural", "political", 
                      "natural", "natural", "natural", 
                      "political", "natural", "natural", 
                      "political"))

# attach to the main table for from- and to-locs
formulas <- formulas %>% 
  left_join(types_groups %>% 
              rename(
                from_type = type_original,
                from_type_group = type_grouped,
                from_type_major = type_major),
            by = "from_type") %>% 
  left_join(types_groups %>% 
              rename(
                to_type = type_original,
                to_type_group = type_grouped,
                to_type_major = type_major),
            by = "to_type") 


rivers_long <- bind_rows(
  rivers |> select(id,latitude = source_lat,    longitude = source_lon),
  rivers |> select(id,latitude = mouth_lat,    longitude = mouth_lon),
  rivers |> select(id,latitude = mid_lat,    longitude = mid_lon)
) %>% arrange(id)

## sample river coordinates

sample_coord_rivers <- function(x=formulas,y=rivers_long) {

rivers_sampled <- y %>% group_by(id) %>% sample_n(1)

## replace from/to river coordinates randomly from the river set
f <- x |>
  left_join(rivers_sampled |> rename(r_from_lat = latitude, r_from_lon = longitude), 
            by = c("from_id" = "id")) |>
  left_join(rivers_sampled |> rename(r_to_lat = latitude, r_to_lon = longitude),   
            by = c("to_id" = "id")) |>
  mutate(
    from_latitude  = coalesce(if_else(from_type == "river", r_from_lat, NA), from_latitude),
    from_longitude = coalesce(if_else(from_type == "river", r_from_lon, NA), from_longitude),
    to_latitude    = coalesce(if_else(to_type   == "river", r_to_lat,   NA), to_latitude),
    to_longitude   = coalesce(if_else(to_type   == "river", r_to_lon,   NA), to_longitude)
  ) |>
  select(-r_from_lat, -r_from_lon, -r_to_lat, -r_to_lon)

##  distances / clock hands etc.

formulas_d <- f %>% 
  filter(lang != "it") %>% 
  rowwise() %>% 
  mutate(dist_haversine = distHaversine(
    c(from_longitude, from_latitude),
    c(to_longitude, to_latitude)) 
      / 1000) %>% 
  ungroup() %>% 
  # make numbers human-readable
  mutate(dist_haversine = round(dist_haversine, digits = 3),
         dist_longer_1k = ifelse(dist_haversine > 1000, "long", "short")) %>% 
  distinct() %>%  # remove duplicated from-to from the same poem
  
  # calculate new coordinates
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude) 

## degrees
formulas_d <- formulas_d %>% 
  mutate(angle_east_deg = atan2(y2_0, x2_0) * 180 / pi,
         angle_north_deg = (90 - angle_east_deg + 360) %% 360) 
return(formulas_d)
}

## gini's 

calculate_borderness_entropy <- function(x) {

for(i in 1:length(x)) {
  
  df <- x[[i]]
  ## circular absolute distance between angles
  circ_dist <- abs(
    (
      (outer(df$angle_east_deg, df$angle_east_deg, "-") + 180) %% 360) - 180
)
  ## only lower
  distances <- circ_dist[lower.tri(circ_dist)]
  
  ## bin values 
  bins <- seq(0, 180, by = 5) 
  hist_vals <- hist(distances, breaks = bins, plot = FALSE)
  
  ## probabilities of a differences over binned 0,180 interval
  p <- hist_vals$counts / sum(hist_vals$counts)
  
  ## gini and entropy
  df <- df %>% mutate(diff_entropy = -sum(p[p > 0] * log(p[p > 0])),
                      diff_gini = ineq(p, type="Gini"))
  
  x[[i]] <- df
  
}
  
formulas_ineq <- x %>%
  bind_rows() %>% 
  group_by(from_id) %>% 
  mutate(n = max(row_number()), # number of times a place appears
         n_langs = length(unique(lang))) %>%  # number of languages referencing the place
  ungroup() %>% 
  select(lang, n_langs,n,starts_with("from_"), starts_with("diff_")) %>%
  filter(!duplicated(from_id)) # table for the model
return(formulas_ineq)
}

### MODEL

fit_model <- function(model=m0m2,x=formulas_ineq) {

fit_rivers <- update(model,newdata=x)

types <- unique(x$from_type_major)

# Build new data frame
newdata <- tibble(
  from_type_major = types,
  n = 10,
  n_langs = 3)


posterior_means <- posterior_epred(fit_rivers, newdata = newdata)


summary_df <- as_tibble(posterior_means)
colnames(summary_df) <- types


# average diff in means ~0.3 entropy points
m <- mean(unlist(summary_df[,1] - summary_df[,2]))

pp_diff <- table((summary_df[,1] - summary_df[,2]) > 0)
# 95% of the time "political" predictions have larger mean
delt <- pp_diff[2]/sum(pp_diff)

return(c(m,delt))
}

```


Alright, so all that reproduces the data. Now we just sample river coordinates, do angle/direction calculations, clock hands, entropy, and refit model with each resampling.

```{r}
### all together

res <- vector(mode="list",length=100)

for(i in 1:100) {
# sample rivers
formulas_d <- sample_coord_rivers(formulas,rivers_long)

# prepare a list of DF holding all "FROM A" directions
df_list <- formulas_d %>%
  filter(from_id %in% from_select$from_id) %>%
  mutate(diff_entropy = 0,diff_gini=0) %>%
  group_by(from_id) %>% group_split()

x<-calculate_borderness_entropy(df_list)
res[[i]] <- fit_model(m0m2,x)

}

res_df <- map(res, ~tibble(effect = .x[1], diff = .x[2])) |> list_rbind()

res_df %>% ggplot() + geom_histogram(aes(effect)) + 
  geom_vline(aes(xintercept=mean(res_df$effect)),linetype=2,color="grey80") +
  labs(title="Distribution of posterior difference between 'natural' and 'political' entropy")


res_df %>% ggplot() + geom_histogram(aes(diff)) + 
  geom_vline(aes(xintercept=mean(res_df$diff)),linetype=2,color="grey80") +
  labs(title="Distribution of posterior differences between 'natural' and 'political' means ")
```


