---
title: "04_borders_model"
format: html
---

## Direction delta's

**!! NB !! will need `formulas_d` dataset`

The idea is to quantify how "structured" are different types of location (are cities really centers? are bodies of water really borders?)

Step 1. Take a location (e.g. Rome). Calculate all *differences* between outgoing directions
Step 2. Measure "spikiness"/"entropy" of a resulting distribution. This is our estimand E.
Step 3. Each location has E: we want to ask what influences the structure: does "borderness" depend on the type of the location.
Step 4. Fit a model: "E ~ type * log(n_occurences) + n_languages"

n_occurences controls for obvious increase in structure with decrease of observations.

```{r}
library(ineq)

# prepare a list of DF holding all "FROM A" directions
df_list <- formulas_d %>%
  filter(from_id %in% from_select$from_id) %>% mutate(diff_entropy = 0,diff_gini=0) %>%  group_by(from_id) %>% group_split()

df_list_to <- formulas_d %>%
  filter(to_id %in% to_select$to_id) %>% mutate(diff_entropy = 0,diff_gini=0) %>%  group_by(to_id) %>% group_split()

for(i in 1:length(df_list)) {
  
  df <- df_list[[i]]
  ## circular absolute distance between angles
  circ_dist <- abs(
    (
      (outer(df$angle_east_deg, df$angle_east_deg, "-") + 180) %% 360) - 180
)
  ## only lower
  distances <- circ_dist[lower.tri(circ_dist)]
  
  ## bin values 
  bins <- seq(0, 180, by = 5) 
  hist_vals <- hist(distances, breaks = bins, plot = FALSE)
  
  ## probabilities of a differences over binned 0,180 interval
  p <- hist_vals$counts / sum(hist_vals$counts)
  
  ## gini and entropy
  df <- df %>% mutate(diff_entropy = -sum(p[p > 0] * log(p[p > 0])),
                      diff_gini = ineq(p, type="Gini"))
  
  df_list[[i]] <- df
  

}


formulas_ineq <- df_list %>%
  bind_rows() %>% 
  group_by(from_id) %>% 
  mutate(n = max(row_number()), # number of times a place appears
         n_langs = length(unique(lang))) %>%  # number of languages referencing the place
  ungroup() %>% 
  select(lang, n_langs,n,starts_with("from_"), starts_with("diff_")) %>%
  filter(!duplicated(from_id)) # table for the model

cor(formulas_ineq$diff_entropy,formulas_ineq$diff_gini)


formulas_to_ineq <- df_list_to %>%
  bind_rows() %>% 
  group_by(to_id) %>% 
  mutate(n = max(row_number()), # number of times a place appears
         n_langs = length(unique(lang))) %>%  # number of languages referencing the place
  ungroup() %>% 
  select(lang, n_langs,n,starts_with("to_"), starts_with("diff_")) %>%
  filter(!duplicated(to_id)) # table for the model


```

```{r}

intersection <- intersect(formulas_ineq$from_placename,formulas_to_ineq$to_placename)
union1 <- union(formulas_ineq$from_placename,formulas_to_ineq$to_placename)

length(intersection)
length(union1)

top_from <- formulas_ineq %>% filter(n>=8)
top_to <- formulas_to_ineq %>% filter(n>=6)

## changes in structure?
top_from %>% 
  left_join(top_to,by = c("from_id" = "to_id")) %>%
  filter(from_placename==to_placename) %>% 
  mutate(delta = diff_entropy.x - diff_entropy.y) %>%
  select(from_placename,delta) %>%
  arrange(-abs(delta)) 
```


### Fitting the model

```{r}
library(brms)

## regularizing priors
priors <- c(
  prior(normal(0, 0.5), class = "b"),
  prior(normal(1, 0.5), class = "Intercept")
)

## prior predictive check (just make sure it's reasonable enough)
fit_prior <- brm(
  diff_entropy ~ from_type_major*log(n) + n_langs,
  data = formulas_ineq,
  family = gaussian(),
  prior = priors,
  sample_prior = "only"
)

pp_check(fit_prior)


m0m2 <- brm(
  diff_entropy ~ from_type_major*log(n) + n_langs,
  data = formulas_ineq,
  family = gaussian(),
  chains = 4,
  cores = 4,
  prior = priors,
  iter = 4000,
  save_model = "m0m2",
)

saveRDS(m0m2,file="../models/entropy_major_type.rds")


m0m2_to <- brm(
  diff_entropy ~ to_type_major*log(n) + n_langs,
  data = formulas_to_ineq,
  family = gaussian(),
  chains = 4,
  cores = 4,
  prior = priors,
  iter = 4000,
  save_model = "m0m2_to",
)

saveRDS(m0m2_to,file="../models/entropy_to_major_type.rds")



```

```{r}
m0m2 <- readRDS("../models/entropy_major_type.rds")
m0m2_to <- readRDS("../models/entropy_to_major_type.rds")
## from
pp_check(m0m2)
conditional_effects(m0m2)

summary(m0m2)

## to
pp_check(m0m2_to)
conditional_effects(m0m2_to)

summary(m0m2_to)

```

### Plots

Redrawing conditional effects here:

```{r}
ce <- conditional_effects(m0m2)
ce_df <- as_tibble(ce[["n:from_type_major"]])



p3 <- ggplot(ce_df,aes(x = n, y = estimate__, color = from_type_major, fill = from_type_major)) +
  geom_line() +
  geom_ribbon(
    aes(ymin = lower__, ymax = upper__),
    alpha = 0.2,
    color = NA
  ) +
  theme_minimal(base_size=12) +
  labs(
    x = "Number of occurrences of a location",
    y = "Predicted entropy"
  ) +
  geom_vline(aes(xintercept=10),linetype=2,color="grey40") + 
  scale_color_manual(values=met.brewer("Cassatt1")[c(2,8)]) +
  scale_fill_manual(values=met.brewer("Cassatt1")[c(2,8)]) + labs(color="Location type",fill="Location type",title="Interaction between type and number of locations")
``` 


```{r}
# Get all levels of type
types <- unique(formulas_ineq$from_type_major)

# Build new data frame
newdata <- tibble(
  from_type_major = types,
  n = 10,
  n_langs = 3)


posterior_draws <- posterior_predict(m0m2, newdata = newdata)
posterior_means <- posterior_epred(m0m2, newdata = newdata)


summary_df <- as_tibble(posterior_means)
colnames(summary_df) <- types


# average diff in means ~0.43 entropy points
mean(unlist(summary_df[,1] - summary_df[,2]))

pp_diff <- table((summary_df[,1] - summary_df[,2]) > 0)
# 96% of the time "political" predictions have larger mean
pp_diff[2]/sum(pp_diff)

means1 <- summary_df %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "pred") %>% 
  group_by(type) %>% summarize(m1=mean(pred))

p1 <- summary_df %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "pred") %>%
  ggplot(aes(pred,reorder(type,-pred),color=type,fill=type)) +
  
  geom_density_ridges(size=0.5,alpha=0.3,scale=1) + 
  geom_vline(data=means1, aes(xintercept=m1,color=type),linetype=2) +
  theme_minimal(base_size = 12) + 
  scale_color_manual(values=met.brewer("Cassatt1")[c(2,8)]) +
  scale_fill_manual(values=met.brewer("Cassatt1")[c(2,8)]) + 
  guides(color="none",fill="none") +
  labs(x="",y=NULL,title="FROM A locations")
```
"To" distances

```{r}
# Get all levels of type
types <- unique(formulas_to_ineq$to_type_major)

# Build new data frame
newdata <- tibble(
  to_type_major = types,
  n = 10,
  n_langs = 3)


posterior_draws <- posterior_predict(m0m2_to, newdata = newdata)
posterior_means <- posterior_epred(m0m2_to, newdata = newdata)


summary_df <- as_tibble(posterior_means)
colnames(summary_df) <- types


# average diff in means ~0.3 entropy points
mean(unlist(summary_df[,1] - summary_df[,2]))

pp_diff <- table((summary_df[,1] - summary_df[,2]) > 0)
# 95% of the time "political" predictions have larger mean
pp_diff[2]/sum(pp_diff)


means1 <- summary_df %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "pred") %>% 
  group_by(type) %>% summarize(m1=mean(pred))

p2 <- summary_df %>% 
  pivot_longer(cols = everything(), names_to = "type", values_to = "pred") %>%
  ggplot(aes(pred,reorder(type,-pred),color=type,fill=type)) +
  
  geom_density_ridges(size=0.5,alpha=0.3,scale=1) + 
  geom_vline(data=means1, aes(xintercept=m1,color=type),linetype=2) +
  theme_minimal(base_size = 12) + 
  labs(x="Predicted marginal entropy",y=NULL,title="TO B locations") +
  scale_color_manual(values=met.brewer("Cassatt1")[c(2,8)]) +
  scale_fill_manual(values=met.brewer("Cassatt1")[c(2,8)]) + labs(color="Location type",fill="Location type") + guides(color="none",fill="none")


library(patchwork)

(p1 / p2) | p3

ggsave("../plots/Fig_model.png",height = 6,width = 10)
```

