---
title: "01 geodata exploration"
format: md
editor: visual
---

# Explore directions of from_to formula

```{r, echo=TRUE, results = 'hide', warning=F, message=F}
library(tidyverse)

# maps pckg
library(sf)
library(ggspatial)
library(rnaturalearth)
library(rnaturalearthdata)

library(ggrepel)
library(MetBrewer)
library(cowplot)
library(kableExtra)
theme_set(theme_minimal())

library(geosphere) # distHavestine()
```

Read data as a table (flatted, merged, and manually cleaned json-s)

```{r}
formulas <- read.csv("../data/formulas_table.csv")
glimpse(formulas)
```

N formulas per language

```{r, include=T, echo=F}
formulas %>% 
  count(lang, sort = T)
```

## Most freq loc vs from_to loc

Look if most most freq locations also appear in from_to formula; calculate kendall rank correlation between the most frequent locations in corpus overall and in formulas

```{r, echo=F}
# load loc freq data
loc_freq <- read.csv("../data/locations.csv")

# glimpse(loc_freq)

loc_freq <- loc_freq %>% 
  # add ranks within groups
  group_by(corpus) %>% 
  mutate(rank = row_number()) %>% 
  ungroup() %>% 
  rename(lang = corpus)

f_long_from <- formulas %>% 
  select(lang, from_id) %>% 
  rename(qid = from_id)

f_long <- f_long_from %>% 
  rbind(formulas %>% select(lang, to_id) %>% rename(qid = to_id))

f_long_t <- f_long %>% 
  group_by(lang) %>% 
  count(qid, sort = T) %>% 
  mutate(rank_ft = row_number()) %>% 
  ungroup() %>% 
  mutate(id = paste0(lang, "_", qid)) %>% 
  select(-lang, -qid) %>% 
  left_join(loc_freq %>% mutate(id = paste0(lang, "_", qid)), by = "id") %>% 
  filter(rank_ft < 21) %>% 
  arrange(lang) %>% 
  select(-lat, -lon) %>% 
  rename(from_to_count = n, 
         total_count = count, 
         rank_total = rank) %>% 
  select(id, qid, lang, label, rank_ft, rank_total, from_to_count, total_count) %>% 
  filter(!is.na(qid))

f_long_t

kcor <- (c(cor(f_long_t %>% filter(lang == "cs") %>% pull(rank_ft), 
      f_long_t %>% filter(lang == "cs") %>% pull(rank_total),
          method = "kendall"),
  cor(f_long_t %>% filter(lang == "de") %>% pull(rank_ft), 
      f_long_t %>% filter(lang == "de") %>% pull(rank_total),
          method = "kendall"),
  cor(f_long_t %>% filter(lang == "en") %>% pull(rank_ft), 
      f_long_t %>% filter(lang == "en") %>% pull(rank_total),
          method = "kendall"),
  cor(f_long_t %>% filter(lang == "fr") %>% pull(rank_ft), 
      f_long_t %>% filter(lang == "fr") %>% pull(rank_total),
          method = "kendall"),
  cor(f_long_t %>% filter(lang == "ru") %>% pull(rank_ft), 
      f_long_t %>% filter(lang == "ru") %>% pull(rank_total),
          method = "kendall"),
  cor(f_long_t %>% filter(lang == "sl") %>% pull(rank_ft), 
      f_long_t %>% filter(lang == "sl") %>% pull(rank_total),
          method = "kendall")))

tibble(lang = c("cs", "de", "en", "fr", "ru", "sl"),
       kendall_cor = round(kcor, 3))

mean(kcor)

rm(f_long, f_long_from, loc_freq, f_long_t, kcor)
```

## Distances

I calculated Haversine (bird/plane flight) distance between the from-\>to coordinates, the distance is in km.

### calculation

```{r, echo=T}
# calculate haversine distances between from_to points in current table

formulas_d <- formulas %>% 
  rowwise() %>% 
  mutate(dist_haversine = distHaversine(
    c(from_longitude, from_latitude),
    c(to_longitude, to_latitude)) 
      / 1000) %>% 
  ungroup() %>% 
  # make numbers human-readable
  mutate(dist_haversine = round(dist_haversine, digits = 3)) 

head(formulas_d %>% select(lang, text, from_placename, to_placename, dist_haversine))
```

# Fig 1: Example

```{r, echo=T}
exmpl <- formulas_d %>% 
   filter(str_detect(to_placename, "Rome")) %>% 
   filter(str_detect(author_name, "Whittier")) # specific line in eng

# exmpl <- formulas_d %>% 
#   filter(str_detect(from_placename, "Rhine") & str_detect(to_placename, "Danube"))

# exmpl
exmpl$text

exmpl$dist_haversine

exmpl %>% select(from_placename, from_longitude, from_latitude, 
                 to_placename, to_longitude, to_latitude)

exmpl %>% 
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude) %>% 
  select(from_placename, from_longitude, from_latitude, 
                 to_placename, to_longitude, to_latitude,
         x1_0, y1_0, x2_0, y2_0)
```

Basic map 1

```{r, include=F, echo=F}
world <- ne_countries(scale = "medium", returnclass = "sf")

europe <- world[which(world$continent == "Europe"),]

# europe <- world

# from malta to rome
map_1 <- ggplot(europe) +
  geom_sf(fill = "#cbeedb",
          alpha = 0.5
          ) +
  coord_sf(xlim = c(9, 18), ylim = c(34, 43), expand = FALSE) + 
  
  geom_curve(data = exmpl,
             aes(x = from_longitude, y = from_latitude,
                xend = to_longitude, yend = to_latitude),
             linewidth = 5, curvature = 0,
             colour = "slategray3", alpha = 0.9) +
  
  # malta label
  geom_text_repel(data = exmpl,
                  aes(x = from_longitude, y = from_latitude, label = from_placename),
                  size = 14, col = met.brewer("Cassatt2")[8], fontface = "bold",
                  vjust = 1.5, 
                  segment.color = "transparent") + 
  
  # rome label
  geom_text_repel(data = exmpl,
                  aes(x = to_longitude, y = to_latitude, label = to_placename),
                  size = 14, col = met.brewer("Cassatt2")[3], fontface = "bold",
                  vjust = 1,  box.padding = 0.7,
                  segment.color = "transparent") + 
  
  # malta point
  geom_point(data = exmpl, aes(x = from_longitude, y = from_latitude), 
             size = 12, alpha = 1, color = met.brewer("Cassatt2")[8]) + 
  
  # rome point
  geom_point(data = exmpl, aes(x = to_longitude, y = to_latitude), 
             size = 12, alpha = 1, color = met.brewer("Cassatt2")[3]) +
  
  
  # annotation_scale(location = "bl", width_hint = 0.4) +
  # annotation_north_arrow(location = "bl", which_north = "true", 
  #       #pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
  #       style = north_arrow_fancy_orienteering)
  
  theme(panel.grid.major = element_line(color = "grey60", 
                                        #linetype = "dashed", 
                                        size = 0.1),
    #panel.background = element_rect(fill = "aliceblue"))
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.background = element_rect(fill = "transparent",
                                       colour = NA_character_),
        panel.background = element_rect(fill = "transparent", 
                                        colour = NA_character_))

map_1

# ggsave(file = "../plots/chr_lt/2_1.png", plot = last_plot(), bg = "transparent",
#        dpi = 300, width = 6.5, height = 6.5) 
```

Map 2: no background line plot

```{r, include=F, echo=F}
map_2 <- exmpl %>% 
  ggplot() + 
  geom_segment(aes(x = from_longitude, xend = to_longitude,
                   y = from_latitude, yend = to_latitude),
               color = "slategray3", linewidth = 4
               ) +
  geom_point(aes( x = from_longitude, y = from_latitude), 
             color = met.brewer("Cassatt2")[8], size = 10) + 
  geom_point(aes( x = to_longitude, y = to_latitude), 
             color = met.brewer("Cassatt2")[3], size = 10) + 
  
  geom_text(aes(x = from_longitude, y = from_latitude),
            label = "FROM(35.88, 14.5)", color = met.brewer("Cassatt2")[8],
            size = 10, vjust = -1, hjust = -0.01,
            fontface = "bold") + 
  geom_text(aes(x = to_longitude, y = to_latitude),
            label = "TO(41.89, 12.48)", color = met.brewer("Cassatt2")[3],
            size = 10, vjust = -1, hjust = -0.01,
            fontface = "bold") + 
  coord_equal() + 
  expand_limits(x = c(14, 23), y = c(35, 44)) + 
  labs(x = "", y = "") + 
  theme(plot.background = element_rect(fill = "transparent",
                                       colour = NA_character_),
        panel.background = element_rect(fill = "transparent", 
                                        colour = NA_character_),
        axis.text = element_blank())

map_2  

# ggsave(file = "../plots/chr_lt/2_2.png", plot = last_plot(), bg = "transparent",
#        dpi = 300, width = 5, height = 5) 
```

Map 3: transposed

```{r, include=F, echo=F}
map_3 <- exmpl %>% 
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude) %>% 
  
  ggplot() + 
  
  annotate("point", x = 0, y = 0, size = 200, 
           alpha = 0.4, 
           fill = "grey10", # met.brewer("Cassatt1")[5], 
           colour = "grey10") + #met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, linewidth = 2, colour = "slategray3") + 
  geom_vline(xintercept = 0, lty = 2, linewidth = 2, colour = "slategray3") +

  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = "slategray2", linewidth = 4) + 
  
  geom_point(aes(x = x1_0, y = y1_0), 
             colour = met.brewer("Cassatt2")[8], size = 10) + 
  geom_point(aes(x = x2_0, y = y2_0), 
             colour = met.brewer("Cassatt2")[3], size = 10) + 
  geom_text(aes(x = x1_0, y = y1_0), 
            colour = met.brewer("Cassatt2")[8], label = "FROM\n(0, 0)",
            vjust = -0.5, size = 10,
            fontface = "bold") + 
  geom_text(aes(x = x2_0, y = y2_0), 
            colour = met.brewer("Cassatt2")[3], label = "TO",
            vjust = -1, size = 10,
            fontface = "bold") + 
  expand_limits(x = c(-5, 4), y = c(-0.5, 8.5)) + 
  coord_equal() + 
  labs(x = "", y = "") + 
  theme(plot.background = element_rect(fill = "transparent",
                                       colour = NA_character_),
        panel.background = element_rect(fill = "transparent", 
                                        colour = NA_character_))
  
map_3

# ggsave(file = "../plots/chr_lt/2_3.png", plot = last_plot(), bg = "transparent",
#        dpi = 300, width = 5, height = 5)
  
```

Map 4: Clock plot:

```{r, include=F, echo=F}
map_4 <- exmpl %>% 
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude) %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 140, 
           alpha = 0.4, 
           fill = "grey10", # met.brewer("Cassatt1")[5], 
           colour = "grey10") + #met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, linewidth = 1.5, colour = "slategray3") + 
  geom_vline(xintercept = 0, lty = 2, linewidth = 1.5, colour = "slategray3") +

  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               arrow = arrow(length = unit(10, "pt")),
               colour = "slategray2", linewidth = 3) + 
  coord_equal() + 
  annotate("point", x = 0, y = 0, size = 7,
           fill = "slategray2", colour = "slategray2") + 
  scale_x_continuous(limits = c(-7, 7)) + 
  scale_y_continuous(limits = c(-7, 7)) + 
  
  # geom_point(aes(x = x1_0, y = y1_0), 
  #            colour = "slategray1", size = 10) + 
  # geom_point(aes(x = x2_0, y = y2_0), 
  #            colour = "pink3", size = 10) + 
  # geom_text(aes(x = x1_0, y = y1_0), 
  #           colour ="slategray1", label = "FROM\n(0, 0)",
  #           vjust = -0.5, size = 10,
  #           fontface = "bold") + 
  # geom_text(aes(x = x2_0, y = y2_0), 
  #           colour ="pink3", label = "TO",
  #           vjust = -1, size = 10,
  #           fontface = "bold") + 
  # expand_limits(x = c(-3, 1), y = c(-1.5, 8)) + 
  labs(x = "", y = "") + 
  theme(plot.background = element_rect(fill = "transparent",
                                       colour = NA_character_),
        panel.background = element_rect(fill = "transparent", 
                                        colour = NA_character_))
  
map_4

# ggsave(file = "../plots/chr_lt/2_4.png", plot = last_plot(), bg = "transparent",
#        dpi = 300, width = 5, height = 5)
```

```{r, eval = F}
plot_grid(map_1, map_2, map_3, map_4,
          nrow = 1,
          labels = c("1", "2", "3", "4"), label_size = 36)
```

```{r, eval=F, echo=F}
ggsave("../plots/Fig_1.png", plot = last_plot(),
       bg = "white", dpi = 300, width = 20, height = 5)
```

```{r, include=F}
rm(exmpl, map_1, map_2, map_3, map_4)
```

# Dist distribution

Poet's mind is often flying not that far away?

```{r, include=T, echo=F}
formulas_d %>% 
  #filter(lang == "sl") %>% 
  ggplot(aes(x = lang, y = dist_haversine, group = lang)) + 
  geom_jitter(alpha = 0.2, width = 0.2) +
  labs(title = "Distribution of all distances",
       y = "Dist Haversine (km)")
```

Most of the distances are actually very small; but for some traditions there are quite a portion of longer ones.

# Fig 2: dist distribution

```{r, include=T, echo=F}
dist_raw <- formulas_d %>% 
  filter(lang != "it") %>% 
  filter(dist_haversine > 0) %>% 
  ggplot(aes(x = lang, y = dist_haversine, group = lang)) + 
  geom_jitter(alpha = 0.2, width = 0.2, colour = met.brewer("Cassatt2")[10]) + 
  geom_violin(fill = met.brewer("Cassatt2")[4], alpha = 0.7) + 
  labs(y = "Distance (km)",
       x = "Corpus") + 
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))

dist_raw
```

```{r, include=T, echo=F}
dist_log <- formulas_d %>% 
  filter(lang != "it") %>% 
  filter(dist_haversine > 0) %>% 
  ggplot(aes(y = factor(lang, levels = c("sl", "ru", "fr", "en", "de", "cs")), 
             x = dist_haversine, group = lang)) + 
  geom_jitter(alpha = 0.2, width = 0.2, colour = met.brewer("Cassatt2")[10]) + 
  geom_violin(fill = met.brewer("Cassatt2")[4], alpha = 0.7) + 
  labs(x = "Distance, km (log scale)",
       y = "Corpus") + 
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) + 
  scale_x_log10()

dist_log
```

```{r, eval=F, include=F, echo=F}
plot_grid(dist_raw, NULL, dist_log,
          rel_widths = c(1, 0.1, 1), nrow = 1,
          labels = c("A", "", "B"), label_size = 16)


ggsave("../plots/Fig_2.png", plot = last_plot(),
       bg = "white", dpi = 300, width = 7, height = 4)
```

```{r, include=F}
rm(dist_log, dist_raw)
```

### dist summary stats

Calculate mean & median dist for each corpus + 3rd quantile

```{r, include=T, echo=F}
dist_summary <- formulas_d %>% 
  group_by(lang) %>% 
  summarise(dist_mean = mean(dist_haversine),
            dist_median = median(dist_haversine),
            third_quant = quantile(dist_haversine, 0.75)) %>% 
  ungroup()

dist_summary
```

Attach means & medians and add grouping lables: if grouped to short/long based on the means in each language, proportion of short/long is almost the same, about 30/70

```{r, include = T, echo = F}

totals <- formulas_d %>% 
  count(lang) %>% 
  rename(n_total = n)

formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_long = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  count(lang, dist_long) %>% 
  left_join(totals, by = "lang") %>% 
  mutate(perc = round((n/n_total)*100, 1))
```

\% longer than 1000 km

```{r, include=T, echo=F}
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_long = ifelse(dist_haversine > 800, "long", "short")) %>% 
  count(lang, dist_long) %>% 
  left_join(totals, by = "lang") %>% 
  mutate(perc = round((n/n_total)*100, 1))
```

The longest distances examples:

```{r, include=T, echo=F}
formulas_d %>% 
  arrange(desc(dist_haversine)) %>% 
  head(10) %>% 
  select(lang, text, from_type, to_type, from_placename, to_placename, dist_haversine)

```

# Types

How distances are related to the types of locations in from-to formula?

### Overview

Total

```{r, include=T, echo=F}
formulas_d %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  count(type_pair, sort = T) %>% 
  head(20)
```

Create larger groups

```{r, echo=T}
unique(c(formulas_d$from_type, formulas_d$to_type))

types_groups <- tibble(type_original = unique(c(formulas_d$from_type, formulas_d$to_type)),
       type_grouped = c("water", "other", "land", 
                        "city", "water", "mountain", 
                        "country", "water", "settlement", 
                        "city", "settlement", "city", 
                        "water", "land", "land", 
                        "land", "mountain", "water", 
                        "settlement", "water", "mountain", 
                        "settlement"), 
       type_major = c("natural", "natural", "natural", 
                      "political", "natural", "natural", 
                      "political", "natural", "political", 
                      "political", "political", "political", 
                      "natural", "natural", "political", 
                      "natural", "natural", "natural", 
                      "political", "natural", "natural", 
                      "political"))

# attach to the main table for from- and to-locs
formulas_d <- formulas_d %>% 
  left_join(types_groups %>% 
              rename(
                from_type = type_original,
                from_type_group = type_grouped,
                from_type_major = type_major),
            by = "from_type") %>% 
  left_join(types_groups %>% 
              rename(
                to_type = type_original,
                to_type_group = type_grouped,
                to_type_major = type_major),
            by = "to_type") 
```

```{r, echo=F}
formulas_d %>% count(from_type_group) # pull(n) %>% sum()
formulas_d %>% count(to_type_group) # %>% pull(n) %>% sum()

formulas_d %>% count(from_type_major) # %>% pull(n) %>% sum()
formulas_d %>% count(to_type_major) # %>% pull(n) %>% sum()
```

```{r, echo=T}
types_groups
```

Count 22 classes

```{r, include=T, echo=F}
x <- formulas_d %>% select(from_type) %>% 
  rename(to_type = from_type)

formulas_d %>% 
  select(to_type) %>% 
  rbind(x) %>% 
  count(to_type, sort = T)
```

```{r, include=T, echo=F}
x <- formulas_d %>% 
  mutate(lang_type = paste0(lang, "_", from_type_major)) %>% 
  count(lang_type)

formulas_d %>% 
  mutate(lang_type = paste0(lang, "_", to_type_major)) %>% 
  count(lang_type) %>% 
  rename(n_x = n) %>% 
  left_join(x, by = "lang_type") %>% 
  mutate(lang = str_extract(lang_type, "^\\w{2}")) %>% 
  left_join(totals, by = "lang") %>% 
  mutate(n_total = n_total * 2, 
         type_total = n_x + n,
         type_perc = round((type_total / n_total) * 100, 1)) %>% 
  select(lang_type, type_total, n_total, type_perc)
```

6 classes distribution

```{r, include=T, echo=F}
x <- formulas_d %>% 
  mutate(lang_type = paste0(lang, "_", from_type_group)) %>% 
  count(lang_type)

formulas_d %>% 
  mutate(lang_type = paste0(lang, "_", to_type_group)) %>% 
  count(lang_type) %>% 
  rename(n_x = n) %>% 
  left_join(x, by = "lang_type") %>% 
  mutate(lang = str_extract(lang_type, "^\\w{2}")) %>% 
  left_join(totals, by = "lang") %>% 
  mutate(n_total = n_total * 2, 
         type_total = n_x + n,
         type_perc = round((type_total / n_total) * 100, 1)) %>% 
  # group_by(lang) %>%
  # summarise(total = sum(type_perc))
  select(lang_type, type_total, n_total, type_perc)
```

### long / short

#### dist ranks

The most frequent types of connection for short and long distances (based on the mean); numbers are ranks

```{r, include=T, echo=F}
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  count(type_pair, dist_type, sort = T) %>% 
  group_by(dist_type) %>% 
  #slice_max(order_by = n, n = 10) %>% 
  mutate(rank = row_number()) %>% 
  select(-n) %>% 
  pivot_wider(names_from = dist_type, values_from = rank) %>% 
  arrange(-desc(short)) %>% head(20)

formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  count(type_pair, dist_type, sort = T) %>% 
  group_by(dist_type) %>% 
  #slice_max(order_by = n, n = 10) %>% 
  mutate(rank = row_number()) %>% 
  select(-n) %>% 
  pivot_wider(names_from = dist_type, values_from = rank) %>% 
  arrange(-desc(long)) %>% head(20)
```

Village –\> village distances

```{r, include=T, echo=F}
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(type_pair == "village --> village") %>% 
  slice_max(dist_haversine, n = 10) %>% 
  select(lang, dist_haversine, text, from_placename, to_placename)

formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(type_pair == "village --> village") %>% 
  pull(dist_haversine) %>% 
  median()
```

### median and mean distances by top-5 patterns

```{r, include = T, echo = F}
# ranks in short distances types
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(dist_type == "short") %>% 
  count(type_pair, sort = T) %>% 
  head(7)

top_short_types <- formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(dist_type == "short") %>% 
  count(type_pair, sort = T) %>% 
  head(7) %>% 
  pull(type_pair)

formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(dist_type == "short") %>% 
  filter(type_pair %in% top_short_types) %>% 
  group_by(type_pair) %>% 
  summarise(mean_dist = mean(dist_haversine),
            med_dist = median(dist_haversine))
```

```{r, include = T, echo = F}
# ranks in long distances types
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(dist_type == "long") %>% 
  count(type_pair, sort = T) %>% 
  head(7)

top_short_types <- formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(dist_type == "long") %>% 
  count(type_pair, sort = T) %>% 
  head(7) %>% 
  pull(type_pair)

formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(dist_type == "long") %>% 
  filter(type_pair %in% top_short_types) %>% 
  group_by(type_pair) %>% 
  summarise(mean_dist = mean(dist_haversine),
            med_dist = median(dist_haversine))
```

#### by language

```{r, include=T, echo=F}
formulas %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  group_by(lang) %>% 
  count(type_pair, sort = T) %>% 
  filter(n > 1) %>% 
  slice_max(order_by = n, n = 10) %>% 
  ungroup() %>% 
  mutate(type_pair = paste0(type_pair, " (", n, ")")) %>% 
  group_by(lang) %>% 
  mutate(top_list = paste0(type_pair, collapse = " <br> ")) %>% 
  select(-n, -type_pair) %>% 
  ungroup() %>% 
  distinct() %>% 
  #pivot_wider(names_from = lang, values_from = top_list) %>% 
  knitr::kable(escape = F)
```

#### longer vs shorter distances

```{r, include=T, echo=F}
formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_type = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
  mutate(type_pair = paste0(from_type, " --> ", to_type)) %>% 
  group_by(lang, dist_type) %>% 
  count(type_pair, sort = T) %>% 
  filter(n > 1) %>% 
  slice_max(order_by = n, n = 5) %>% 
  ungroup() 
```

### city parts

```{r, include=T, echo=F}
formulas_d %>% 
  filter(str_detect(from_type, "city part")) %>% 
  count(lang, from_placename, sort = T)

formulas_d %>% 
  filter(str_detect(to_type, "city part")) %>% 
  count(lang, to_placename, sort = T)

# traditions with most city parts present?
x <- formulas_d %>% 
  filter(str_detect(from_type, "city part")) %>% 
  select(lang, from_placename) %>% 
  rename(placename = from_placename)

formulas_d %>% 
  filter(str_detect(to_type, "city part")) %>% 
  select(lang, to_placename) %>% 
  rename(placename = to_placename) %>% 
  rbind(x) %>% 
  group_by(lang) %>% 
  count(placename, sort = T) %>% 
  # filter(n > 1) %>% 
  ungroup() %>% 
  arrange(-desc(lang))

formulas_d %>% 
  mutate(from_to_pair = paste0(from_type, " --> ", to_type)) %>% 
  filter(from_to_pair == "city part --> city part") %>% 
  select(lang, text, from_placename, to_placename)

formulas_d %>% 
  filter(lang == "fr") %>% 
  filter(from_placename == "Moscow Kremlin" | to_placename == "Moscow Kremlin")
```

## Places frequency

### from-to formulas

Just lists of real places, I'm not sure we need to think about this now (not too many observations also)

```{r, include=T, echo=F}
formulas %>% 
  mutate(from_to_pair = paste0(from_placename, " --> ", to_placename)) %>% 
  count(from_to_pair, sort = T) %>% 
  head(20)
```

By language (n \> 1)

```{r, include=T, echo=F}
formulas %>% 
  mutate(from_to_pair = paste0(from_placename, " --> ", to_placename)) %>% 
  group_by(lang) %>% 
  count(from_to_pair, sort = T) %>% 
  slice_max(order_by = n, n = 5, with_ties = F) %>% 
  filter(n > 1) %>% 
  knitr::kable()
```

```{r, eval = F, echo = F}
formulas_d %>% 
  filter(lang == "fr") %>% 
  filter(from_placename == "Moscow Kremlin" | to_placename == "Moscow Kremlin")
```

### most from / to places

Filter n \> 3 !

```{r, include=T, echo=F}
formulas %>% 
  filter(lang != "it") %>% 
  group_by(lang) %>% 
  count(from_placename, from_type, sort = T) %>% 
  slice_max(order_by = n, n = 5) %>% 
  filter(n > 3)
```

```{r, include=T, echo=F}
formulas %>% 
  filter(lang != "it") %>% 
  group_by(lang) %>% 
  count(to_placename, to_type, sort = T) %>% 
  slice_max(order_by = n, n = 5) %>% 
  filter(n > 3)
```

### most freq places in general

General frequencies (highly dependent on N formulas from a language)

```{r, include=T, echo=F}
x <- formulas %>% 
  select(from_id, from_placename, from_type) %>% 
  rename(id = from_id,
         placename = from_placename,
         type = from_type)

places_freq <- formulas %>% 
  select(to_id, to_placename, to_type) %>% 
  rename(id = to_id, 
         placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  count(id, placename, sort = T) %>% 
  mutate(rank_freq = row_number())

head(places_freq, 20)
```

Locations mentioned in the most of corpora (either from or to)

```{r}
x <- formulas_d %>% 
  select(from_id, from_placename, from_type, lang) %>% 
  rename(id = from_id,
         placename = from_placename,
         type = from_type) 

freq_ranks <- formulas_d %>% 
  select(to_id, to_placename, to_type, lang) %>% 
  rename(id = to_id, 
         placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  distinct() %>% 
  count(id, placename, type, sort = T) %>% 
  rename(n_corpora = n) %>% 
  filter(n_corpora > 3) %>% 
  arrange(-desc(placename)) %>% 
  arrange(desc(n_corpora)) %>% 
  # attach abs freq
  left_join(places_freq %>% select(id, n), by = "id")

freq_ranks
```
