---
title: "02 clock plots"
format: md
editor: visual
---

# Direction Analysis and Compass Plots

This notebook generates compass plots and analyses directional patterns in the formulas. To fully reproduce Figures 3 and 4, load the direction plots from the `03_directions_plot.qmd` into the R environment.

-   Calculate Coordinates: Transforms 'from' coordinates to (0,0) and adjusts 'to' coordinates accordingly;

-   Long Distances: Creates six compass plots showing long distances directions across corpora;

-   Short Distances: Produces same kind of plots for short-distance formulas;

-   Figure 5: Prepares background map and compasses used in the Fig. 5.

-   Additional - Rivers: Generate compass plots for additional river coordinates to examine their directional distribution.

```{r, warning=F, include=F}
library(tidyverse)
library(geosphere)

library(cowplot)
library(MetBrewer)
theme_set(theme_minimal())

# maps
library(sf)
library(ggspatial)

library(rnaturalearth)
library(rnaturalearthdata)
library(ggrepel)
```

Load data & calculate distances

```{r, include=T, echo=F}
formulas <- read.csv("../data/formulas_table.csv")
# glimpse(formulas)

formulas_d <- formulas %>% 
  rowwise() %>% 
  mutate(dist_haversine = distHaversine(
    c(from_longitude, from_latitude),
    c(to_longitude, to_latitude)) 
      / 1000) %>% 
  ungroup() %>% 
  # make numbers human-readable
  mutate(dist_haversine = round(dist_haversine, digits = 3),
         dist_longer_1k = ifelse(dist_haversine > 1000, "long", "short")) 

sample_n(formulas_d %>% select(lang, text, from_placename, to_placename, dist_haversine), 5)

# add mean dist for long/short grouping
dist_summary <- formulas_d %>% 
  group_by(lang) %>% 
  summarise(mean_dist = mean(dist_haversine),
            .groups = "drop")

# final cleaning & additions
formulas_d <- formulas_d %>% 
  left_join(dist_summary, by = "lang") %>% 
  mutate(dist_longer_mean = ifelse(dist_haversine > mean_dist, "long", "short")) 
```

# Calculate coordinates

```{r}
formulas_nc <- formulas_d %>% 
  # calculate new coordinates
  mutate(x1_0 = 0.0,
         y1_0 = 0.0,
         
         x2_0 = to_longitude - from_longitude,
         y2_0 = to_latitude - from_latitude)

glimpse(formulas_nc)
```

# Long distances

## Fig. 3

```{r, eval=F, echo=F}
lang_labels <- tibble(
  lang = c("cs", "de", "en", "fr", "ru", "sl"),
  lang_full = c("Czech", "German", "English", "French", "Russian", "Slovenian")
)
```

```{r, echo=F}
p_main <- formulas_nc %>% 
  filter(dist_longer_mean == "long") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[5], 
           colour = met.brewer("Cassatt1")[5]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(7, "pt")),
               linewidth = 1,
               alpha = 0.5) + 
  facet_wrap(~lang, ncol = 3) + 
  scale_x_continuous(limits = c(-230, 230)) + 
  scale_y_continuous(limits = c(-230, 230)) + 
  coord_equal() + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 20, face = "bold"),
        # spacing for slides
        panel.spacing = unit(1, "lines"))

p_main
```

```{r, eval=F, inlude=F, echo=F}
# ggsave("../plots/Fig_2.png", plot = last_plot(), 
#        width = 12, height = 10, bg = "white", dpi = 300)


### NB! LOAD PLOT FROM SCR 03

# to save together with the pizza-plot:
# load the plot from scr 03 !

plot_grid(p_main, NA, long_dist_pizza,
          nrow = 1, rel_widths = c(1, 0.1, 1),
          labels = c("A", "", "B"), label_size = 36)

ggsave("../plots/paper/Fig_3.png", plot = last_plot(),
       width = 20, height = 7, bg = "white", dpi = 300)
```

# short distances

```{r, echo=F, results='hide'}

## 1000 km threshold

p_short <- formulas_nc %>% 
  filter(dist_longer_1k == "short") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 70, 
           alpha = 0.6, 
           fill = met.brewer("Cassatt1")[4], 
           colour = met.brewer("Cassatt1")[4]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(4, "pt")),
               alpha = 0.5) + 
  # facet_wrap(~lang, ncol = 1) + 
  coord_equal() + 
  
  scale_x_continuous(limits = c(-15, 15)) + 
  scale_y_continuous(limits = c(-15, 15)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        #axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 14, face = "bold")) + 
  labs(title = "Short distances (<1000 km)")
  
p_short + facet_wrap(~lang, ncol = 3)

p1 <- p_short + facet_wrap(~lang, ncol = 1)
```

```{r, eval=F, inlude=F, echo=F}
ggsave("../plots/01_compass_dist_short.png", plot = last_plot(), 
       width = 10, height = 8, bg = "white", dpi = 300)
```

## Fig. 4: short dist based on mean

```{r, echo=F}
p_short <- formulas_nc %>% 
  filter(dist_longer_mean == "short") %>% 
  
  ggplot() + 
  annotate("point", x = 0, y = 0, size = 100, 
           alpha = 0.4, 
           fill = met.brewer("Cassatt2")[4], 
           colour = met.brewer("Cassatt2")[4]) + 
  
  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_0,
                   y = y1_0, yend = y2_0), 
               colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(4, "pt")),
               linewidth = 1,
               alpha = 0.5) + 
  coord_equal() + 
  
  # facet_wrap(~lang, ncol = 1) + 
  scale_x_continuous(limits = c(-33, 33)) + 
  scale_y_continuous(limits = c(-33, 33)) + 
  
  annotate("point", x = 0, y = 0, size = 1,
           fill = "white", colour = "white") + 
  
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 14, face = "bold")) + 
  # labs(title = "Short distances (dist < mean_dist_lang)") + 
  facet_wrap(~lang, ncol = 3)

p_short
```

```{r, eval = F, echo = F, include = F}
# to save together with the pizza-plot:
# load the plot from scr 03 !

plot_grid(p_short, NA, short_dist_pizza,
          nrow = 1, rel_widths = c(1, 0.1, 1),
          labels = c("A", "", "B"), label_size = 36)

ggsave("../plots/paper/Fig_4.png", plot = last_plot(),
       width = 20, height = 7, bg = "white", dpi = 300)
```

# Fig. 5: most common places

Most common places, map background & transparent compases

```{r, echo=F}
x <- formulas_nc %>% 
  select(from_placename, from_type, lang) %>% 
  rename(placename = from_placename,
         type = from_type) 

freq_ranks <- formulas_nc %>% 
  select(to_placename, to_type, lang) %>% 
  rename(placename = to_placename,
         type = to_type) %>% 
  rbind(x) %>% 
  distinct() %>% 
  count(placename, type, sort = T) %>% 
  rename(n_corpora = n) %>% 
  filter(n_corpora > 3) # %>% 
  # filter(placename %in% c(
  #   "Baltic Sea", "Tagus River", "Paris", "Rome", "Athens", 
  #   "Nile", "Euphrates", "Babylon", "Bethlehem", 
  #   "Dardanelles", "Don", "Moscow"
  # )) 

freq_ranks
```

```{r, include = F, echo = F}
# get locations data for the map
# select only FROM locations for the viz
loc_coord <- formulas_nc %>% 
  filter(from_placename %in% freq_ranks$placename) 

freq_coord <- loc_coord %>% 
  left_join(freq_ranks %>% 
              mutate(grouping = ifelse(placename %in% c("Paris", "Rome", "Vienna"),
                                                    "center", "border")) %>% 
              select(placename, grouping) %>% 
              rename(from_placename = placename), 
            by = "from_placename") # %>% 
  # left_join(freq_ranks %>% select(placename, grouping) %>% 
  #             rename(to_placename = placename,
  #                    grouping_t = grouping)) %>% 
  # mutate(grouping = ifelse(is.na(grouping), grouping_t, grouping)) %>% 
  # select(-grouping_t)

# freq_coord
```

## map

Locations mentioned in 4 or 5 corpora

```{r, echo = F}
# load map
world <- ne_countries(scale = "medium", returnclass = "sf")

# europe <- world[which(world$continent == "Europe"),]


x <- freq_coord %>% 
  select(from_id, from_placename, from_latitude, from_longitude, lang) %>% 
  rename(id = from_id,
         placename = from_placename,
         lat = from_latitude,
         long = from_longitude) 


pl_coord <- freq_coord %>% 
  select(to_id, to_placename, to_latitude, to_longitude, lang) %>% 
  rename(id = to_id,
         placename = to_placename,
         lat = to_latitude,
         long = to_longitude) %>% 
  rbind(x) %>% 
  select(-lang) %>% 
  distinct() %>% 
  filter(placename %in% freq_ranks$placename) 

ggplot(world) +
  geom_sf(#fill = "#cbeedb"
          ) +
  coord_sf(xlim = c(-15, 80), ylim = c(27,62), expand = FALSE) + 
  
  geom_point(data = pl_coord, aes(x = long, y = lat), 
             size = 2, alpha = 0.9, color = "violetred") + 
  geom_text_repel(data = pl_coord, 
                  aes(x = long, y = lat, label = placename),
                  size = 4, color = "midnightblue")

```

```{r, eval=F}
# helper map 

ggplot(world) +
  geom_sf(fill = "cornsilk1",
          color = "cornsilk4",
          linewidth = 1
          ) +
  coord_sf(xlim = c(-15, 68), ylim = c(27,62), expand = FALSE) + 
  
  geom_point(data = pl_coord, aes(x = long, y = lat),
             size = 1, alpha = 0.9, color = "violetred4") + 
  theme(panel.grid.major = element_line(color = "grey60", 
                                        #linetype = "dashed", 
                                        linewidth = 0.1), 
        panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(color = "grey60"))  
  

# ggsave(file = "../plots/3.png", plot = last_plot(), 
#        bg = "white",
#        dpi = 300, width = 20, height = 10)
```

## compasses

Compasses for the places appeared in 4 or 5 corpora

```{r, eval=F, echo=F}
glimpse(freq_coord)
```

```{r, eval = F, echo = F}
freq_coord %>% 
  mutate(base_place = ifelse(from_placename %in% freq_ranks$placename, 
                             from_placename, to_placename)) %>% 
  count(base_place, sort = T)
```

```{r, echo=T}

fixed_length <- 100

freq_coord %>% 
  # filter only from places
  mutate(base_place = ifelse(from_placename %in% freq_ranks$placename, 
                             from_placename, "")) %>% 
  filter(base_place != "") %>% 
  
  # distances (clock handle) length normalization:
  # calculate vector length (hypotenuse) & normalize it
  mutate(
    # hypotenuse
    vec_len = sqrt(x2_0^2 + y2_0^2),
    
    # norm
    x2_norm = (x2_0 / vec_len) * fixed_length,
    y2_norm = (y2_0 / vec_len) * fixed_length
         ) %>% 
  
  # plot
  ggplot() + 
  
  annotate("point", x = 0, y = 0, size = 120,
           alpha = 0.4,
           fill = met.brewer("Cassatt1")[5],
           colour = met.brewer("Cassatt1")[5]) +

  geom_hline(yintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  geom_vline(xintercept = 0, lty = 2, colour = "white", 
             linewidth = 1) + 
  
  geom_segment(aes(x = x1_0, xend = x2_norm,
                   y = y1_0, yend = y2_norm,
                   color = grouping), 
               #colour = met.brewer("Cassatt2")[9],
               arrow = arrow(length = unit(7, "pt")),
               size = 2,
               alpha = 1) + 
  coord_equal() + 
  facet_wrap(~base_place, ncol = 3) + 
  scale_color_manual(values = c(met.brewer("Cassatt2")[9], 
                                met.brewer("Cassatt1")[1])) +
  scale_x_continuous(limits = c(-120, 120)) + 
  scale_y_continuous(limits = c(-120, 120)) + 
  
  annotate("point", x = 0, y = 0, size = 7,
           fill = "violetred4", colour = "violetred4") + 
  
  theme(# plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(), 
        axis.text = element_blank(), 
        axis.title = element_blank(),
        strip.text.x = element_text(size = 14, face = "bold"), 
        
        plot.background = element_rect(fill = "transparent",
                                       colour = NA_character_),
        panel.background = element_rect(fill = "transparent", 
                                        colour = NA_character_))
  
# ggsave(file = "../plots/3_1.png", plot = last_plot(),
#        bg = "transparent",
#        dpi = 300, width = 15, height = 39)
```

------------------------------------------------------------------------

# Additional - Rivers

Read the file with additional coordinates for rivers

```{r, echo=F}
rivers <- read_csv("../data/rivers_with_coords.csv")
```

```{r, eval=F}
glimpse(rivers)

glimpse(formulas_d)
```

Create 3 different sets: with rivers coords for source, mouths, or mid

```{r, echo=T}
upd_coord <- function(rivers_subset, formulas_df) {
  # attach new long and lat
  formulas_upd <- formulas_df %>% 
    left_join(rivers_subset %>% rename(from_id = id), by = "from_id") %>% 
    mutate(from_latitude = ifelse(!is.na(latitude),
                                  latitude, from_latitude),
           from_longitude = ifelse(!is.na(longitude),
                                  longitude, from_longitude)) %>% 
    select(-c(latitude, longitude)) %>% 
    left_join(rivers_subset %>% rename(to_id = id), by = "to_id") %>% 
    mutate(to_latitude = ifelse(!is.na(latitude),
                                latitude, to_latitude),
           to_longitude = ifelse(!is.na(longitude),
                                 longitude, to_longitude)) %>% 
    select(-c(latitude, longitude))
  
  # recalc centralized points
  formulas_upd <- formulas_upd %>% 
    mutate(x1_0 = 0.0,
           y1_0 = 0.0,
           
           x2_0 = to_longitude - from_longitude,
           y2_0 = to_latitude - from_latitude)

  return(formulas_upd)
}


# only sources

# for all rivers get only sources
rivers_source <- rivers %>% 
  select(id, source_lat, source_lon) %>% 
  rename(latitude = source_lat,
         longitude = source_lon)

fr_sources <- upd_coord(rivers_source, formulas_d)

# glimpse(fr_sources)

# only mouths coordinates

# for all rivers get only sources
rivers_mouth <- rivers %>% 
  select(id, mouth_lat, mouth_lon) %>% 
  rename(latitude = mouth_lat,
         longitude = mouth_lon)

fr_mouths <- upd_coord(rivers_mouth, formulas_d)

# glimpse(fr_mouths)

# only mid-point coordinates

# for all rivers get only sources
rivers_mid <- rivers %>% 
  select(id, mid_lat, mid_lon) %>% 
  rename(latitude = mid_lat,
         longitude = mid_lon)

fr_mid <- upd_coord(rivers_mid, formulas_d)
```

Build the clock plots again

```{r, echo=F}

# this function created compass plots for long distances, the same as for Fig. 3; rivers are coloured in red

clock_long <- function(formulas_nc) {
  p_main <- formulas_nc %>% 
    mutate(arrow_col = ifelse(from_type == "river" | to_type == "river",
                              "river", "not_river")) %>% 
    filter(dist_longer_mean == "long") %>% 
    
    ggplot() + 
    annotate("point", x = 0, y = 0, size = 100, 
             alpha = 0.6, 
             fill = met.brewer("Cassatt1")[5], 
             colour = met.brewer("Cassatt1")[5]) + 
    
    geom_hline(yintercept = 0, lty = 2, colour = "white", 
               linewidth = 1) + 
    geom_vline(xintercept = 0, lty = 2, colour = "white", 
               linewidth = 1) + 
    
    geom_segment(aes(x = x1_0, xend = x2_0,
                     y = y1_0, yend = y2_0,
                     colour = arrow_col), 
                 #colour = met.brewer("Cassatt2")[9],
                 arrow = arrow(length = unit(7, "pt")),
                 linewidth = 1,
                 alpha = 0.5) + 
    facet_wrap(~lang, ncol = 3) + 
    scale_x_continuous(limits = c(-230, 230)) + 
    scale_y_continuous(limits = c(-230, 230)) + 
    scale_colour_manual(values = c(met.brewer("Cassatt2")[9],
                                   met.brewer("Cassatt1")[1])) +
    coord_equal() + 
    
    annotate("point", x = 0, y = 0, size = 1,
             fill = "white", colour = "white") + 
    
    theme(plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(), 
          axis.text = element_blank(), 
          axis.title = element_blank(),
          strip.text.x = element_text(size = 20, face = "bold"),
          # spacing for slides
          panel.spacing = unit(1, "lines"))

  return(p_main)
}


# this function created compass plots for short distances, the same as for Fig. 4; rivers are coloured in red

clock_short <- function(formulas_nc) {
  p_short <- formulas_nc %>% 
    # arrow colour col
    mutate(arrow_col = ifelse(from_type == "river" | to_type == "river",
                              "river", "not_river")) %>% 
    filter(dist_longer_mean == "short") %>% 
    
    ggplot() + 
    annotate("point", x = 0, y = 0, size = 100, 
             alpha = 0.4, 
             fill = met.brewer("Cassatt2")[4], 
             colour = met.brewer("Cassatt2")[4]) + 
    
    geom_hline(yintercept = 0, lty = 2, colour = "white", 
               linewidth = 1) + 
    geom_vline(xintercept = 0, lty = 2, colour = "white", 
               linewidth = 1) + 
    
    geom_segment(aes(x = x1_0, xend = x2_0,
                     y = y1_0, yend = y2_0,
                     colour = arrow_col), 
                 #colour = met.brewer("Cassatt2")[9],
                 arrow = arrow(length = unit(4, "pt")),
                 linewidth = 1,
                 alpha = 0.5) + 
    
    
    scale_colour_manual(values = c(met.brewer("Cassatt2")[9],
                                   met.brewer("Cassatt1")[1])) +
    
    coord_equal() + 
    
    # facet_wrap(~lang, ncol = 1) + 
    scale_x_continuous(limits = c(-33, 33)) + 
    scale_y_continuous(limits = c(-33, 33)) + 
    
    annotate("point", x = 0, y = 0, size = 1,
             fill = "white", colour = "white") + 
    
    theme(plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(), 
          axis.text = element_blank(), 
          axis.title = element_blank(),
          strip.text.x = element_text(size = 14, face = "bold")) + 
    # labs(title = "Short distances (dist < mean_dist_lang)") + 
    facet_wrap(~lang, ncol = 3)
  
  return(p_short)
}


```

```{r}
# the function creates the plot for directions distribution (long distances), same as in 03_directions_plot.qmd; it first assigns each formula to a segment, and then calculate proportion of formulas in each segment.

long_dist_pizza <- function(formulas_d) {

  formulas_d <- formulas_d %>% 
    mutate(angle_east_deg = atan2(y2_0, x2_0) * 180 / pi,
           angle_north_deg = (90 - angle_east_deg + 360) %% 360) %>% 
    mutate(
      compass = case_when(
        angle_north_deg < 22.5 | angle_north_deg >= 337.5 ~ "N",
        angle_north_deg < 67.5  ~ "NE",
        angle_north_deg < 112.5 ~ "E",
        angle_north_deg < 157.5 ~ "SE",
        angle_north_deg < 202.5 ~ "S",
        angle_north_deg < 247.5 ~ "SW",
        angle_north_deg < 292.5 ~ "W",
        TRUE                    ~ "NW"
      )
    ) %>% 
    mutate(
      compass = factor(
        compass,
        levels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW")
      )
    )
  
  # percentages
  dist_mean <- formulas_d %>% 
    group_by(lang) %>% 
    summarise(dist_mean = mean(dist_haversine),
              .groups = "drop")
  
  totals <- formulas_d %>% 
    left_join(dist_mean, by = "lang") %>% 
    mutate(dist_longer_mean = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
    count(lang, dist_longer_mean) %>% 
    mutate(lang_d = paste0(lang, "_", dist_longer_mean)) %>% 
    rename(total_f = n)
  
  x <- formulas_d %>% 
    left_join(dist_mean, by = "lang") %>% 
    mutate(dist_longer_mean = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
    mutate(lang_d = paste0(lang, "_", dist_longer_mean)) %>% 
    group_by(lang_d) %>% 
    count(compass) %>% 
    left_join(totals, by = "lang_d") %>% 
    mutate(perc = (n / total_f)) %>% 
    ungroup() %>% 
    mutate(value = 1)  
    
  
  long_dist_pizza_p <- x %>% 
    filter(dist_longer_mean == "long") %>% 
    ggplot(aes(x = compass, y = value)) + 
      geom_col(width = 1,
             position = "fill", colour = "white",
             fill = met.brewer("Cassatt1")[8],
             aes(alpha = perc)) + 
      facet_wrap(~lang) + 
      labs(alpha = "%") + 
      geom_hline(yintercept = 1.0, colour = met.brewer("Cassatt1")[8], alpha = 0.5) + 
      theme(
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        strip.text = element_text(face = "bold", size = 20),
        #legend.position = "bottom",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.text.x = element_text(face = "bold", size = 12)
        ) + 
      coord_polar(start = -pi/8) 
  
  return(long_dist_pizza_p)
}
```

```{r}
# the function creates the plot for directions distribution (short distancs), same as in 03_directions_plot.qmd; it first assign each formula to a segment, and then calculate proportion of formulas in each segment.

short_dist_pizza <- function(formulas_d) {

  formulas_d <- formulas_d %>% 
    mutate(angle_east_deg = atan2(y2_0, x2_0) * 180 / pi,
           angle_north_deg = (90 - angle_east_deg + 360) %% 360) %>% 
    mutate(
      compass = case_when(
        angle_north_deg < 22.5 | angle_north_deg >= 337.5 ~ "N",
        angle_north_deg < 67.5  ~ "NE",
        angle_north_deg < 112.5 ~ "E",
        angle_north_deg < 157.5 ~ "SE",
        angle_north_deg < 202.5 ~ "S",
        angle_north_deg < 247.5 ~ "SW",
        angle_north_deg < 292.5 ~ "W",
        TRUE                    ~ "NW"
      )
    ) %>% 
    mutate(
      compass = factor(
        compass,
        levels = c("N", "NE", "E", "SE", "S", "SW", "W", "NW")
      )
    )
  
  # percentages
  dist_mean <- formulas_d %>% 
    group_by(lang) %>% 
    summarise(dist_mean = mean(dist_haversine),
              .groups = "drop")
  
  totals <- formulas_d %>% 
    left_join(dist_mean, by = "lang") %>% 
    mutate(dist_longer_mean = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
    count(lang, dist_longer_mean) %>% 
    mutate(lang_d = paste0(lang, "_", dist_longer_mean)) %>% 
    rename(total_f = n)
  
  x <- formulas_d %>% 
    left_join(dist_mean, by = "lang") %>% 
    mutate(dist_longer_mean = ifelse(dist_haversine > dist_mean, "long", "short")) %>% 
    mutate(lang_d = paste0(lang, "_", dist_longer_mean)) %>% 
    group_by(lang_d) %>% 
    count(compass) %>% 
    left_join(totals, by = "lang_d") %>% 
    mutate(perc = (n / total_f)) %>% 
    ungroup() %>% 
    mutate(value = 1)  
    
  
  short_dist_pizza_p <- x %>% 
    filter(dist_longer_mean == "short") %>% 
    ggplot(aes(x = compass, y = value)) + 
    geom_col(width = 1,
             position = "fill", colour = "white",
             fill = met.brewer("Cassatt2")[3],
             aes(alpha = perc)) + 
    facet_wrap(~lang) + 
    labs(alpha = "%") +
    geom_hline(yintercept = 1.0, colour = met.brewer("Cassatt2")[3], alpha = 0.5) + 
    theme(
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        strip.text = element_text(face = "bold", size = 20),
        #legend.position = "bottom",
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.text.x = element_text(face = "bold", size = 12)
        ) + 
    coord_polar(start = -pi/8) 
  
  return(short_dist_pizza_p)
}
```

```{r, echo=T, warning=F, message=F}
################## sources

# glimpse(fr_sources)

# long
c_long <- clock_long(fr_sources) + ggtitle("Long dist; river sources") + 
  theme(legend.position = "None")
p_long <- long_dist_pizza(fr_sources) + 
  theme(legend.position = "None")


plot_grid(c_long, NA, p_long,
          nrow = 1, rel_widths = c(1, 0.1, 1),
          labels = c("A", "", "B"), label_size = 36)

# ggsave("../plots/rivers/long_sources.png", plot = last_plot(),
#        width = 20, height = 7, bg = "white", dpi = 300)

# short
c_short <- clock_short(fr_sources) + ggtitle("Short dist; river sources") + 
  theme(legend.position = "None")
p_short <- short_dist_pizza(fr_sources) + 
  theme(legend.position = "None")


plot_grid(c_short, NA, p_short,
          nrow = 1, rel_widths = c(1, 0.1, 1),
          labels = c("A", "", "B"), label_size = 36)

# ggsave("../plots/rivers/short_sources.png", plot = last_plot(),
#        width = 20, height = 7, bg = "white", dpi = 300)

###################### mouths

# glimpse(fr_mouths)

# long
c_long <- clock_long(fr_mouths) + ggtitle("Long dist; river mouths") + 
  theme(legend.position = "None")
p_long <- long_dist_pizza(fr_mouths)  + 
  theme(legend.position = "None")


plot_grid(c_long, NA, p_long,
          nrow = 1, rel_widths = c(1, 0.1, 1),
          labels = c("A", "", "B"), label_size = 36)

# ggsave("../plots/rivers/long_mouths.png", plot = last_plot(),
#        width = 20, height = 7, bg = "white", dpi = 300)

# short
c_short <- clock_short(fr_mouths) + ggtitle("Short dist; river mouths") + 
  theme(legend.position = "None")
p_short <- short_dist_pizza(fr_mouths) + 
  theme(legend.position = "None")


plot_grid(c_short, NA, p_short,
          nrow = 1, rel_widths = c(1, 0.1, 1),
          labels = c("A", "", "B"), label_size = 36)

# ggsave("../plots/rivers/short_mouths.png", plot = last_plot(),
#        width = 20, height = 7, bg = "white", dpi = 300)

################################ mid points
# glimpse(fr_mid)

# long
c_long <- clock_long(fr_mid) + ggtitle("Long dist; river mid-points") + 
  theme(legend.position = "None")
p_long <- long_dist_pizza(fr_mid) + 
  theme(legend.position = "None")


plot_grid(c_long, NA, p_long,
          nrow = 1, rel_widths = c(1, 0.1, 1),
          labels = c("A", "", "B"), label_size = 36)

# ggsave("../plots/rivers/long_mid.png", plot = last_plot(),
#        width = 20, height = 7, bg = "white", dpi = 300)

# short
c_short <- clock_short(fr_mid) + ggtitle("Short dist; river mid-points") + 
  theme(legend.position = "None")
p_short <- short_dist_pizza(fr_mid) + 
  theme(legend.position = "None")


plot_grid(c_short, NA, p_short,
          nrow = 1, rel_widths = c(1, 0.1, 1),
          labels = c("A", "", "B"), label_size = 36)

# ggsave("../plots/rivers/short_mid.png", plot = last_plot(),
#        width = 20, height = 7, bg = "white", dpi = 300)
```
